---
title: "TEAP_data_processing"
author: "Benjamin D. Peterson"
date: "2023-05-23"
output: html_document
---

```{r setup, include = FALSE}
rm(list = ls())
library(ggtree)
library(tidyverse)
library(treeio)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(echo = TRUE)
new.wd <- gsub("\\/[[:alpha:]]+\\/[[:alpha:]]+$", "", getwd())
knitr::opts_knit$set(root.dir = new.wd)
rm(new.wd)
#knitr::opts_knit$set(root.dir = "/Users/benjaminpeterson/Documents/research/BLiMMP")
```

First, I read in the metadata for both MGs and MTs:
  
```{r MG_MT_metadata, include = FALSE}
MG_metadata <- read.csv("metadata/metagenome_metadata.csv") %>%
  rename(omicID = metagenomeID) %>%
  mutate(seqType = "MG") %>%
  select(omicID, startDate, depth, seqType)
MT_metadata <- read.csv("metadata/metatranscriptome_metadata.csv") %>%
  rename(omicID = metatranscriptomeID) %>%
  mutate(seqType = "MT") %>%
  select(omicID, startDate, depth, seqType)
omic_metadata <- rbind(MG_metadata,
                       MT_metadata)
rm(MG_metadata, MT_metadata)
NFs.expression <- readRDS("dataEdited/metatranscriptomes/normalization_vector.rds")
```

# NarG sequences

### Confirm sequences are *NarG*

First I plotted the tree that was generated.

```{r NarG_tree, fig.height = 12}
narG_tree <- read.newick("dataEdited/ABA/narG/narG.tree")
narG_tree_rooted <- phangorn::midpoint(narG_tree)
rm(narG_tree)
ggtree(narG_tree_rooted) +
  geom_tiplab(size = 2) +
  geom_text2(aes(subset = !isTip, label = node),
             size = 1.5,
             hjust = 1,
             vjust = -1) +
  geom_treescale(x = 9)
```

Many of the identified sequences are actually cytoplasmic NxrA sequences.
I removed those.

```{r remove_nxrA_seqs}
narG_tree_subset <- tree_subset(narG_tree_rooted,
                                node = 176,
                                levels_back = 0)
rm(narG_tree_rooted)
ggtree(narG_tree_subset) +
  geom_tiplab(size = 2) +
  geom_text2(aes(subset = !isTip, label = node),
             size = 1.5,
             hjust = 1,
             vjust = -1) +
  geom_treescale(x = 9)
```

### Cluster NarG sequences

Next, I looked at the clustering data to determine a good cut-off to use.

```{r narg_clstr_data_check }
narg_seq_list <- grep(pattern = "^BLI",
                      narG_tree_subset$tip.label,
                      value = TRUE)
narg_clstr_data <- read.table("dataEdited/ABA/narG/narG_cluster_data.tsv",
                              header = TRUE) %>%
  filter(id %in% narg_seq_list)
plot(y = gsub("%", "", narg_clstr_data$clstr_iden) %>% as.numeric() %>% sort(decreasing = TRUE),
     x = 1:length(narg_clstr_data$clstr_iden),
     xlab = "",
     ylab = "Percent ID",
     pch = 16)
abline(h = 97, col = "red")
abline(h = 99, col = "black")
rm(narg_clstr_data, narG_tree_subset)
```

Tough one to pick a cutoff for, but 97 vs. 99 is only a difference of two based on this dataset.
Let's go with 99%.
I went back into the *narG* processing sheet and reran cd-hit with the 0.99 cutoff.
Then I brought that data into R.

### Aggregate and inspect NarG data
```{r narg_clstr_data }
narg_data <- read.table("dataEdited/ABA/narG/narG_derep_99.tsv",
                              header = TRUE) %>%
  filter(geneID %in% narg_seq_list) %>%
  mutate(geneName = "narG") %>%
  select(geneID, geneName, clstr_ID, length) %>%
  mutate(contigID = gsub('_[[:digit:]]+$', '', x = geneID))
rm(narg_seq_list)
```

Add in the metagenome and metatranscriptome data as well:

```{r narG_coverage_data, warning = FALSE, message = FALSE}
normalized.coverage.vector <- readRDS("dataEdited/scg_coverage/scg_normalization_vector.rds")
narG_MG <- read.table("dataEdited/ABA/narG/narG_MG_coverage.tsv",
                      header = FALSE,
                      col.names = c("contigID", "coverage", "contigLength", "metagenomeID")) %>%
  mutate(NF = normalized.coverage.vector[metagenomeID],
         coverage_normalized = coverage*NF) %>%
  select(contigID, metagenomeID, coverage_normalized) %>%
  spread(key = metagenomeID, value = coverage_normalized)
narG_MT <- read.table("dataEdited/ABA/narG/narG_MT_coverage.tsv",
                        header = TRUE) %>%
  mutate(counts_per_bp = counts/effective_length,
         million_counts_per_liter = counts_per_bp*NFs.expression[mtID]/1000000) %>%
  select(geneID, mtID, million_counts_per_liter) %>%
  spread(key = mtID,
         value = million_counts_per_liter)
narg_data <- left_join(narg_data, narG_MG) %>%
  left_join(narG_MT)
rm(narG_MT, narG_MG)
```

Inspect the metagenome and metatranscriptome abundance.

```{r check_narG_data, fig.height=10, echo = FALSE, warning=FALSE, message = FALSE}
plot_data <- narg_data %>%
  gather(key = omicID,
         value = coverage,
         -c(1:5)) %>%
  group_by(clstr_ID, omicID) %>%
  summarise(coverage = mean(coverage)) %>%
  ungroup() %>%
  group_by(omicID) %>%
  summarise(coverage = sum(coverage)) %>%
  ungroup() %>%
  left_join(omic_metadata) %>%
  mutate(date_depth = paste(startDate, "-", depth, "m",
                            sep = "")) 
  
narG_MG <- plot_data %>%
  filter(seqType == "MG") %>%
  ggplot(aes(x = date_depth,
             y = coverage)) +
  geom_point() +
  facet_wrap(~seqType,
             nrow = 2) +
  theme_bw() +
  ylim(c(0, 10)) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust=1)) +
  xlab("")
narG_MT <- plot_data %>%
  filter(seqType == "MT") %>%
  ggplot(aes(x = date_depth,
             y = log(coverage, 10))) +
  geom_point() +
  facet_wrap(~seqType,
             nrow = 2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust=1)) +
  xlab("")
ggpubr::ggarrange(narG_MG,
                  narG_MT,
                  ncol = 1,
                  nrow = 2)
rm(plot_data, narG_MT, narG_MG)
```







# DsrA sequences

### Identify reductive vs. oxidative *dsrA* genes

```{r DsrA_tree_root_node_ID, fig.height = 40, include = FALSE}
dsrA_tree <- read.newick("dataEdited/ABA/dsrA/dsrA.tree")
ggtree(dsrA_tree) +
  geom_tiplab(size = 2) +
  geom_text2(aes(subset = !isTip, label = node),
             size = 2.5,
             hjust = 1,
             vjust = -1) +
  geom_treescale(x = 9)
```

The first step was to look at the tree that I generated and identify the rooting node.
The node leading to the DsrB seqs is 738, so we'll root on that.
Next, I needed to split up the sequences by reductive vs. oxidative DsrA sequences.
To do this, I looked at the rooted tree with the nodes labeled:
  
```{r DsrA_tree_rooted, fig.height = 40}
dsrA_tree_rooted <- root(phy = dsrA_tree,
                         node = 738)
ggtree(dsrA_tree_rooted) +
  geom_tiplab(size = 2) +
  geom_treescale(x = 9) +
  geom_text2(aes(subset = !isTip, label = node),
             size = 2.5,
             hjust = 1,
             vjust = -1)
rm(dsrA_tree)
```

The branch of oxidative DsrA sequences is a subtree of the whole tree, so I pulled out those sequence IDs to mark as oxidative (node 476).
Quick check to make sure this is right:
  
```{r oxi_DsrA_tree, fig.height = 10}
oxi_dsrA_tree_rooted <- tree_subset(dsrA_tree_rooted,
                                    node = 476,
                                    levels_back = 0)
ggtree(oxi_dsrA_tree_rooted) +
  geom_tiplab(size = 2) +
  geom_treescale(x = 9) +
  geom_text2(aes(subset = !isTip, label = node),
             size = 2.5,
             hjust = 1,
             vjust = -1)
```

This looks correct.
Next, I classified all the sequences from this study in that tree as oxidative, and the remaining DsrA seqs as reductive.

```{r classify_dsrA_seqs}
dsrA_seqs <- grep("BLI2[0-1]_",
                  dsrA_tree_rooted$tip.label,
                  value = TRUE)
oxi_dsrA_seqs <- grep("BLI2[0-1]_",
                      oxi_dsrA_tree_rooted$tip.label,
                      value = TRUE)
red_dsrA_seqs <- dsrA_seqs[-which(dsrA_seqs %in% oxi_dsrA_seqs)]
dsrA_type_vector <- c(rep("oxi_dsrA", length(oxi_dsrA_seqs)),
                      rep("red_dsrA", length(red_dsrA_seqs)))
names(dsrA_type_vector) <- c(oxi_dsrA_seqs, red_dsrA_seqs)
rm(oxi_dsrA_seqs, red_dsrA_seqs, dsrA_seqs, oxi_dsrA_tree_rooted, dsrA_tree_rooted)
```


### Cluster sequences

Next I looked at the clustering values that we used to determine a clustering cutoff:
  
```{r dsrA_clstr_data_check, echo = FALSE}
dsrA_clstr_data <- read.table("dataEdited/ABA/dsrA/dsrA_cluster_data.tsv",
                              header = TRUE)
plot(y = gsub("%", "", dsrA_clstr_data$clstr_iden) %>% as.numeric() %>% sort(decreasing = TRUE),
     x = 1:length(dsrA_clstr_data$clstr_iden),
     xlab = "",
     ylab = "Percent ID",
     pch = 16)
abline(h = 97, col = "red")
rm(dsrA_clstr_data)
```

Using a cutoff of 0.97 here will work, so I re-clustered everything with CD-HIT at that 97% cutoff:

```{zsh engine.opts='-i'}
conda activate bioinformatics
cd-hit -g 1 \
        -i dataEdited/ABA/dsrA/dsrA.faa \
        -o dataEdited/ABA/dsrA/dsrA_clustered.faa \
        -c 0.97 \
        -n 5 \
        -d 0
python ../../programs/HomeBio/bin/FM_CDHIT_parsing.py --clstr_in dataEdited/ABA/dsrA/dsrA_clustered.faa.clstr \
                                                  --clstr_out dataEdited/ABA/dsrA/dsrA_clustered_97.tsv
rm dataEdited/ABA/dsrA/dsrA_clustered.faa dataEdited/ABA/dsrA/dsrA_clustered.faa.clstr
```


### Aggregate and inspect dsrA data

Now that all of the sequences have been clustered, I read them in and aggregated the cluster info with the MG and MT data.

```{r dsrA_clstr_data }
dsrA_data <- read.table("dataEdited/ABA/dsrA/dsrA_clustered_97.tsv",
                              header = TRUE) %>%
  mutate(geneName = dsrA_type_vector[geneID]) %>%
  select(geneID, geneName, clstr_ID, length) %>%
  mutate(contigID = gsub('_[[:digit:]]+$', '', x = geneID))
rm(dsrA_type_vector)

dsrA_MG <- read.table("dataEdited/ABA/dsrA/dsrA_MG_coverage.tsv",
                      header = FALSE,
                      col.names = c("contigID", "coverage", "contigLength", "metagenomeID")) %>%
  mutate(NF = normalized.coverage.vector[metagenomeID],
         coverage_normalized = coverage*NF) %>%
  select(contigID, metagenomeID, coverage_normalized) %>%
  spread(key = metagenomeID, value = coverage_normalized)
dsrA_MT <- read.table("dataEdited/ABA/dsrA/dsrA_MT_coverage.tsv",
                        header = TRUE) %>%
  mutate(counts_per_bp = counts/effective_length,
         million_counts_per_liter = counts_per_bp*NFs.expression[mtID]/1000000) %>%
  select(geneID, mtID, million_counts_per_liter) %>%
  spread(key = mtID,
         value = million_counts_per_liter)
dsrA_data <- left_join(dsrA_data, dsrA_MG) %>%
  left_join(dsrA_MT)
rm(dsrA_MG, dsrA_MT)
```


```{r check_red_dsrA_data, fig.height=10, echo = FALSE, warning=FALSE, message = FALSE}
plot_data_red_dsrA <- dsrA_data %>%
  filter(geneName == "red_dsrA") %>%
  gather(key = omicID,
         value = coverage,
         -c(1:5)) %>%
  group_by(clstr_ID, omicID) %>%
  summarise(coverage = mean(coverage)) %>%
  ungroup() %>%
  group_by(omicID) %>%
  summarise(coverage = sum(coverage)) %>%
  ungroup() %>%
  left_join(omic_metadata) %>%
  mutate(date_depth = paste(startDate, "-", depth, "m",
                            sep = "")) 
  
dsrA_MG <- plot_data_red_dsrA %>%
  filter(seqType == "MG") %>%
  ggplot(aes(x = date_depth,
             y = coverage)) +
  geom_point() +
  facet_wrap(~seqType,
             nrow = 2) +
  theme_bw() +
  ylim(c(0, 5)) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust=1)) +
  xlab("")
dsrA_MT <- plot_data_red_dsrA %>%
  filter(seqType == "MT") %>%
  ggplot(aes(x = date_depth,
             y = coverage)) +
  geom_point() +
  facet_wrap(~seqType,
             nrow = 2) +
  theme_bw() +
  # ylim(c(0, 4)) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust=1)) +
  xlab("")
ggpubr::ggarrange(dsrA_MG,
                  dsrA_MT,
                  ncol = 1,
                  nrow = 2)
rm(plot_data_red_dsrA, dsrA_MG, dsrA_MT)
```

Look at oxidative DsrA as well:

```{r check_oxi_dsrA_data, fig.height=10, echo = FALSE, warning=FALSE, message = FALSE}
plot_data_oxi_dsrA <- dsrA_data %>%
  filter(geneName == "oxi_dsrA") %>%
  gather(key = omicID,
         value = coverage,
         -c(1:5)) %>%
  group_by(clstr_ID, omicID) %>%
  summarise(coverage = mean(coverage)) %>%
  ungroup() %>%
  group_by(omicID) %>%
  summarise(coverage = sum(coverage)) %>%
  ungroup() %>%
  left_join(omic_metadata) %>%
  mutate(date_depth = paste(startDate, "-", depth, "m",
                            sep = "")) 
  
dsrA_MG <- plot_data_oxi_dsrA %>%
  filter(seqType == "MG") %>%
  ggplot(aes(x = date_depth,
             y = coverage)) +
  geom_point() +
  facet_wrap(~seqType,
             nrow = 2) +
  theme_bw() +
  ylim(c(0, 25)) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust=1)) +
  xlab("")
dsrA_MT <- plot_data_oxi_dsrA %>%
  filter(seqType == "MT") %>%
  ggplot(aes(x = date_depth,
             y = coverage)) +
  geom_point() +
  facet_wrap(~seqType,
             nrow = 2) +
  theme_bw() +
  ylim(c(0, 550)) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust=1)) +
  xlab("")
ggpubr::ggarrange(dsrA_MG,
                  dsrA_MT,
                  ncol = 1,
                  nrow = 2)
rm(plot_data_oxi_dsrA, dsrA_MG, dsrA_MT)
```




# Aggregate all TEAP data

```{r aggregate_TEAP_data}
TEAP_data <- rbind(narg_data,
                   dsrA_data) %>%
  mutate(clstr_ID = paste(geneName, "-",
                          stringr::str_pad(string = clstr_ID,
                                           side = "left",
                                           pad = "0",
                                           width = 3),
                          sep = ""))
write.csv(x = TEAP_data,
          file = "dataFinal/TEAP_data.csv",
          quote = FALSE,
          row.names = FALSE)
```

