---
title: "2019 MeHg production assays"
author: "Benjamin D. Peterson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(knitr)
library(tidyverse)

cb.translator <- readRDS("/Users/benjaminpeterson/Box/ancillary_science_stuff/colors/colorblind_friendly_colors_R/colorblind_friendly_colors.rds")
```


In this document, I'll go through the data we collected from the 2019 MeHg production assays.
In here, I'm mostly checking to see how well we did with the assays, and what we might need to focus on going forward.


### BLiMMP_trip_003: Take 1

Let's take a look at the first set of incubations.
We did not collect an Exo profile due to equipment issues. 
We only got geochemical data from this profile, and only got sulfide and FC data in addition to the Hg data. 
We focused on collecting one sample from just below the oxic-anoxic interface, and one from much deeper. 
The deeper sample smelled of sulfide. 
Let's check out the raw Hg data first.

``` {r MeHg BLiMMP_trip_003, echo=FALSE, out.width = "50%", out.height = "50%", message = FALSE}
# Read in a PDF of the document
graph.path <- "/Users/benjaminpeterson/Documents/research/BLiMMP/results/incubations/methylation/incubations2019/BLiMMP_trip_003_incubations.png"
knitr::include_graphics(graph.path)

```

Also looked at it as percent MeHg of total Hg pool.

``` {r fraction MeHg BLiMMP_trip_003, echo=FALSE, out.width = "50%", out.height = "50%", message = FALSE}
# Read in a PDF of the document
graph.path <- "/Users/benjaminpeterson/Documents/research/BLiMMP/results/incubations/methylation/incubations2019/BLiMMP_003_per_Me198Hg.png"
knitr::include_graphics(graph.path)

```
Don't really change how it looks.

Let's also visualize as the change in MeHg from t0 to t1.
I calculated this as the excess Me^198^Hg at t=1 minus the excess Me^198^Hg at t=0, with no adjustment for samples below DL, then normalized that value to 24 hours of production, since many of these only incubated for 22 hours or so.
Need to think more about if I should be accounting for DL here.

``` {r change_in_MeHg_BLiMMP_trip_003, echo=FALSE, message = FALSE}
library(patchwork)
# Read in a PDF of the document
data.path <- "/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/incubations/MeHg/incubations2019_Me198Hg_production.csv"
MeHg.data <- read.csv(data.path,
                      stringsAsFactors = FALSE) %>%
  filter(tripID == "BLiMMP_trip_003") %>%
  select(incubationID, depth, treatment, Me198Hg_production)
p1 <- MeHg.data %>%
  filter(depth == 10.3) %>%
  ggplot(aes(x = treatment,
             y = Me198Hg_production)) +
  geom_jitter(position = position_jitter(0.1)) +
  ylim(min(MeHg.data$Me198Hg_production)-0.001,
       max(MeHg.data$Me198Hg_production)+0.003) +
  theme_classic()
p2 <- MeHg.data %>%
  filter(depth == 20) %>%
  ggplot(aes(x = treatment,
             y = Me198Hg_production)) +
  geom_jitter(position = position_jitter(0.1)) + 
  ylim(min(MeHg.data$Me198Hg_production)-0.001,
       max(MeHg.data$Me198Hg_production)+0.003) +
  theme_classic()
p1 + p2

```

We see that there appears to be no Me^198^Hg production at 10.3m.
Obviously, there's still no production of Me^198^Hg when we filter it.
However, at 20m, we see appreciable amounts of MeHg being produced.
Two of the incubations (BLiMMP_MA_0022 and 0023) produce approximately 0.03ng/L of MeHg in 24 hours, but the third incubation (BLiMMP_MA_0021) was substantially lower.
For this incubation, I had noted in my field notes that I did not mix this bag well between injecting the tracer and retrieving the t0 sample.
Thus, I likely withdrew most of the tracer before it had time to mix in.
Let's check out the THg data to see if this is the case.

``` {r check total Hg from BLiMMP_003, echo=FALSE, message=FALSE}
Hg.data <- read.csv("/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/incubations/THg/incubations2019_THg.csv", stringsAsFactors = FALSE) %>%
  filter(tripID == "BLiMMP_trip_003",
         depth == 20) %>%
  select(incubationID, t, THg_198) %>%
  mutate(t = paste("THg at ",
                   t,
                   sep = "")) %>%
  spread(t, THg_198) %>%
  left_join(MeHg.data %>%
              select(incubationID, Me198Hg_production)) %>%
  arrange(incubationID)
kable(Hg.data, caption = "THg levels in BLiMMPP_003 incubations at 20m")

```

We can see that the initial THg count is much higher, suggesting that we did in fact not mix things in properly.
However, it's not that much higher and the final THg value is still pretty high, so unlikely that the lower MeHg production is completely due to pulling out all the ^198^Hg during the t=0 sample.

Additionally, I noticed that the sulfide levels in this bag at the end of the incubation were substantially lower than in the other bags from this depth.
I doubt this is related to withdrawing the spike early, more likely just some O~2~ intrusion into the bag, which could also account for the lower Me^198^Hg production. 

```{r sulfide_data_incubations_BLiMMP_trip_003,  echo=FALSE, message = FALSE, fig.height = 5, fig.width = 5}

sulfide.data <- read.csv("/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/waterChemistry/sulfide/MA_data.csv",
                         stringsAsFactors = FALSE) %>%
  filter(tripID == "BLiMMP_trip_003")

# Set up color vector
color.vector <- c("green", "red")
names(color.vector) <- c("unfiltered-unamended", "filtered-unamended")
# Set up depth constraints
y_max <- ceiling(max(c(sulfide.data$S_conc_uM, sulfide.data$S_conc_uM)))
y_min <- floor(min(c(sulfide.data$S_conc_uM, sulfide.data$S_conc_uM)))

par(mar = c(4.5, 4.5, 3, 1),
    mfrow = c(1, 1))
invisible(
  stripchart(S_conc_uM ~ depth,
            data = sulfide.data,
            vertical = TRUE,
            cex = 1,
            ylim = c(y_min, y_max),
            method = "jitter",
            pch = 18,
            main = "Sulfide concentration from\nincubations from BLiMMP_trip_003",
            xlab = "Depth of incubation",
            ylab = "Sulfide concentration (uM)",
            cex.main = 0.9))
  
  sulfide.data.deep <- sulfide.data %>%
    filter(depth == 20)
  text(x = 1.75,
       y = sulfide.data.deep$S_conc_uM,
       labels = sulfide.data.deep$incubationID,
       cex = 0.7)



```


Again, I wanted to look at this as a fraction of total Hg pool.
So, here we have the change in %MeHg from t0 to t1.

``` {r change in fraction MeHg BLiMMP_trip_003, echo=FALSE, out.width = "50%", out.height = "50%", message = FALSE}
# Read in a PDF of the document
graph.path <- "/Users/benjaminpeterson/Documents/research/BLiMMP/results/incubations/methylation/incubations2019/BLiMMP_003_change_in_per_Me198Hg.png"
knitr::include_graphics(graph.path)

```

Doesn't change our interpretation much, if at all.

#### Takeaways

1. We can consistently detect MeHg production in Mendota.
2. Methylation seems to only occur at depths where sulfide accumulation has started (preliminary).



### BLiMMP_trip_005: Second round of incubations

In the second round of incubations, we added an additional depth, as well as a molybdate treatment.
The uppermost depth (11.7m) was taken from the small turbidity spike right below the oxic-anoxic interface. 
Sulfide was quite low at this point, only 5.4µM.
The next depth was 14m, where sulfide was up to 18.3µM.
The final depth was 20.1m, where sulfide had accumulated to relatively high levels (80µM). 
Leucine uptake rates were highest at the sample from the oxic-anoxic interface, approximately twice as high as they were at the other two depths.
Let's look at the Hg data.

```{r MeHg BLiMMP_trip_005, echo = FALSE, out.height= "50%", out.width="60%"}

# Read in the PNG file with the incubation data
graph.path <- "/Users/benjaminpeterson/Documents/research/BLiMMP/results/incubations/methylation/incubations2019/BLiMMP_trip_005_incubations.png"
include_graphics(graph.path)
```

Again, there was no detectable methylation under any condition in the uppermost layer, where there was no sulfide accumulation.
The two samples in the lower sections of the water column showed no production of MeHg in the filtered incubations.
They showed approximately similar levels of MeHg production in the unfiltered, unamended sample, but the data was pretty noisy, with both having one replicate that showed little to no MeHg production.
We'll look at those replicates later.
Finally, when we added molybdate, the MeHg production shut down again, to pretty much the same levels we saw in the filtered controls.

As before, here it is as fraction MeHg:

```{r fraction MeHg BLiMMP_trip_005, echo = FALSE, out.height= "50%", out.width="60%"}

# Read in the PNG file with the incubation data
graph.path <- "/Users/benjaminpeterson/Documents/research/BLiMMP/results/incubations/methylation/incubations2019/BLiMMP_005_per_Me198Hg.png"
include_graphics(graph.path)
```

Pretty much the same.


Below I plotted this up as the change in Me^198^Hg from t0 to t1.

``` {r change_in_MeHg_BLiMMP_trip_005, fig.width = 8, fig.height = 4, echo=FALSE, message = FALSE}
library(patchwork)

# Generate needed color vector
color.vector <- c(cb.translator["vermillion"],
                  cb.translator["bluishgreen"],
                  cb.translator["black"])
names(color.vector) <- c("filtered\nunamended",
                         "unfiltered\nunamended",
                         "unfiltered\nmolybdate")

# Read in a PDF of the document
data.path <- "/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/incubations/MeHg/incubations2019_Me198Hg_production.csv"
MeHg.data <- read.csv(data.path,
                      stringsAsFactors = FALSE) %>%
  filter(tripID == "BLiMMP_trip_005") %>%
  select(incubationID, depth, treatment, Me198Hg_production)
plot.list <- list()
for (depth.of.interest in unique(MeHg.data$depth)) {
  depth.of.interest.m <- paste(depth.of.interest, "m", sep = "")
  plot.list[[depth.of.interest.m]] <- MeHg.data %>%
    filter(depth == depth.of.interest) %>%
    mutate(treatment = gsub("-", "\n", treatment)) %>%
    mutate(treatment = factor(treatment, levels =  c("filtered\nunamended",
                                                     "unfiltered\nunamended",
                                                     "unfiltered\nmolybdate"))) %>%
    ggplot(aes(x = treatment,
               y = Me198Hg_production,
               color = treatment)) +
    geom_jitter(position = position_jitter(0.1)) +
    scale_color_manual(values = color.vector) +
    theme_classic() +
    theme(legend.position = "none") +
    ylim(min(MeHg.data$Me198Hg_production)-0.001,
         max(MeHg.data$Me198Hg_production)+0.003) +
    xlab("") +
    ggtitle(depth.of.interest.m)
    
}

plot.list[[1]] + plot.list[[2]] + plot.list[[3]]

```

Viewing the data this way tells pretty much the same story.

And finally, viewing it as change in %MeHg:

```{r change in fraction MeHg BLiMMP_trip_005, echo = FALSE, out.height= "50%", out.width="60%"}

# Read in the PNG file with the incubation data
graph.path <- "/Users/benjaminpeterson/Documents/research/BLiMMP/results/incubations/methylation/incubations2019/BLiMMP_005_change_in_per_Me198Hg.png"
include_graphics(graph.path)
```



Now, let's look at those anamolous incubations with low Me^198^Hg production.
I compared the MeHg production to the amount of sulfide left in the bag at the end of the incubation to see if it explained anything.

```{r sulfide_data_incubations_BLiMMP_trip_005,  echo=FALSE, message = FALSE, fig.height = 4, fig.width = 8}

names(color.vector) <- c("filtered-unamended",
                         "unfiltered-unamended",
                         "unfiltered-molybdate")
par(mfrow = c(1, 2),
    mar = c(4.5, 4.5, 3, 1))

sulfide.data <- read.csv("/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/waterChemistry/sulfide/MA_data.csv",
                         stringsAsFactors = FALSE) %>%
  filter(tripID == "BLiMMP_trip_005",
         depth == 14) %>%
  left_join(MeHg.data) %>%
  select(incubationID, treatment, S_conc_uM, Me198Hg_production)

WC.S.levels <- read.csv("/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/waterChemistry/sulfide/WC_data.csv",
                                   stringsAsFactors = FALSE) %>%
            filter(tripID == "BLiMMP_trip_005",
                   depth == 14) %>%
            select(S_conc_uM) %>%
            unlist(use.names = FALSE)
anomalous.point <- sulfide.data %>%
  filter(incubationID == "BLiMMP_MA_0042")

plot(x = jitter(sulfide.data$S_conc_uM),
     y = jitter(sulfide.data$Me198Hg_production),
     col = color.vector[sulfide.data$treatment],
     pch = 18,
     ylim = c(-0.01, max(sulfide.data$Me198Hg_production)*1.2),
               xlim = c(0, max(sulfide.data$S_conc_uM)*1.2),
               xlab = "Sulfide concentration\nat end of incubation(uM)",
               ylab = "MeHg production after 24hr (ng/L/day)",
               main = paste("MeHg production compared\nto sulfide levels at 14m\nWater column = ",
                            WC.S.levels,
                            "µM",
                            sep = ""),
               cex.main = 0.85)
text(x = anomalous.point$S_conc_uM + 1.5,
     y = anomalous.point$Me198Hg_production - 0.005,
    labels = anomalous.point$incubationID,
    col = color.vector[anomalous.point$treatment])


sulfide.data <- read.csv("/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/waterChemistry/sulfide/MA_data.csv",
                         stringsAsFactors = FALSE) %>%
  filter(tripID == "BLiMMP_trip_005",
         depth == 20.1) %>%
  left_join(MeHg.data) %>%
  select(incubationID, treatment, S_conc_uM, Me198Hg_production)

WC.S.levels <- read.csv("/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/waterChemistry/sulfide/WC_data.csv",
                                   stringsAsFactors = FALSE) %>%
            filter(tripID == "BLiMMP_trip_005",
                   depth == 20.1) %>%
            select(S_conc_uM) %>%
            unlist(use.names = FALSE)
anomalous.point <- sulfide.data %>%
  filter(incubationID == "BLiMMP_MA_0051")
plot(x = jitter(sulfide.data$S_conc_uM),
     y = jitter(sulfide.data$Me198Hg_production),
     col = color.vector[sulfide.data$treatment],
     pch = 18,
     ylim = c(-0.01, max(sulfide.data$Me198Hg_production)*1.2),
               xlim = c(0, max(sulfide.data$S_conc_uM)*1.2),
               xlab = "Sulfide concentration\nat end of incubation(uM)",
               ylab = "MeHg production after 24hr (ng/L/day)",
               main = paste("MeHg production compared\nto sulfide levels at 19.3m\nWater column = ",
                            WC.S.levels,
                            "µM",
                            sep = ""),
               cex.main = 0.85)
text(x = anomalous.point$S_conc_uM - 18,
     y = anomalous.point$Me198Hg_production + 0.002,
    labels = anomalous.point$incubationID,
    col = color.vector[anomalous.point$treatment])

```

We'll start with 14m, on the left.
BLiMMP_MA_0042 has low production of Me^198^Hg relative to the other two replicates. 
Here we might see an explanation for why that is.
BLiMMP_MA_0042 also had no sulfide left in that bag at the end of it, whereas the two bags with noticeable MeHg production did have sulfide remaining.
This could indicate a major shift in the redox state.
I suspect that there may have been substantial O~2~ leakage into the bags, which would lead to the oxidation of sulfide to sulfate. 
It seems that we may have had a similar issue with the filtered controls and the molybdate-injected bags, as one of the filtered controls and two of the molybdate-amended samples have no remaining sulfide, and other filtered control with less than 1µM.
However, the samples where there is some sulfide also show much lower levels of MeHg production than we see in the two unfiltered-unamended samples with sulfate present, suggesting that there is no MeHg production under these conditions.

Next we'll look at the graph on the right, which is the same thing, but at 20m deep.
Here, BLiMMP_MA_0051 is the unfiltered and unamended incubation with the low production of Me^198^Hg.
However, the sulfide in this sample isn't particularly low, suggesting that there wasn't an infusion of O~2~ into this bag.
Could be due to issue with overall levels of Hg in the bags, which I address below.

One general point of concern is that even in the bags with sulfide remaining, the sulfide levels are substantially lower than what was in the water column when we originally collected it.
If this is leakage of O~2~ into the bags, the sulfide may be acting as a redox buffer, consuming the O~2~, preserving the microbial community and allowing it to continue methylating.
This would complicate our finding of no methylation occurring in the non-sulfidic incubations, which is obviously problematic.
I think I can test for this though.
If it is O~2~ leakage, we would expect the oxidized sulfide to have been converted to sulfate.
Once I can get back in the lab, I'll analyze these samples for sulfate, to see if the sulfide has indeed been oxidized to sulfate.

Another potential reason the MeHg production in BLiMMP_MA_0051 is so low would be if we had lost a bunch of the overall ^198^Hg pool.
So, I looked at this below.

``` {r check total Hg from BLiMMP_005, echo=FALSE, message=FALSE}
Hg.data <- read.csv("/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/incubations/THg/incubations2019_THg.csv", stringsAsFactors = FALSE) %>%
  filter(tripID == "BLiMMP_trip_005",
         depth == 20.1) %>%
  select(incubationID, t, treatment, THg_198) %>%
  mutate(t = paste("THg at ",
                   t,
                   sep = "")) %>%
  spread(t, THg_198) %>%
  left_join(MeHg.data %>%
              select(incubationID, Me198Hg_production)) %>%
  arrange(incubationID)
kable(Hg.data, caption = "THg levels in BLiMMPP_003 incubations at 20m")

```

While the data is very noisy, and we do see a trend in decreased T^198^Hg over the course of the experiment, BLiMMP_MA_0051 does not stand out as having particularly low ^198^Hg.
In fact, it is nearly identical to BLiMMP_MA_0052.
Thus, this doesn't explain it either.
I also have flow cytometry samples for this bag, which could explain changes in the microbial community.
If we get a chance to analyze that, we'll look to that for answers, otherwise we'll just have to leave it at what it.

#### Takeaways

1. Once again, we only see biotic MeHg production in the regions of the water column with detectable sulfide accumulation.
2. Errant data can sometimes be explained by the ancillary data we're collecting (BLiMMP_MA_0042), but not necessarily always (BLiMMP_MA_0051).
3. Molybdate amendment strongly inhibits the production of MeHg.


### BLiMMP_trip_006: Final round for 2019

In our third and final round of incubations for 2019, we went back to just doing 2 depths.
We also reduced the number of replicates for our different conditions to 2, to try to increase the number of amendments we could try, as a prelude to the 2020 season.
We stuck with the usual set of filtered controls, ambient conditions, and molybdate amendments.
However, we also added starch amendments, algal biomass amendments, and starch + molybdate.
The logic behind this was that we had seen a large number of fermentative hgcA+ bacteria in our genomes, and we hypothesized that they would be able to drive MeHg production independent of sulfate reduction.
First we needed to see if they would increase MeHg production, however, so that's why we started here.
We did do the starch-molybdate amendment to try to get a start on this though.

The first depth was just under the oxic/anoxic interface.
Normally we'd see the turbidity spike here, but the turbidity spike was right on the oxic/anoxic interface, so we went about 30-50cm below that, where DO was at 0.
This put us at about 16.2m.
Surprisingly, sulfide was already almost 60µM here.
Mn was also slightly enriched here relative to the deeper sample, up over 8µM (about 0.46ppm), suggesting localized Mn cycling at the oxic/anoxic interface.
The second depth was at 19.3m.
Here, sulfide was nearly 100µM, but Mn was under 6µM.
Let's look at the Hg data.

```{r MeHg BLiMMP_trip_006, echo = FALSE, out.height= "75%", out.width="75%"}

# Read in a PDF of the document
graph.path <- "/Users/benjaminpeterson/Documents/research/BLiMMP/results/incubations/methylation/incubations2019/BLiMMP_trip_006_incubations.png"
include_graphics(graph.path)
```

Here it as percent MeHg of total Hg pool:

```{r fraction MeHg BLiMMP_trip_006, echo = FALSE, out.height= "75%", out.width="75%"}

# Read in a PDF of the document
graph.path <- "/Users/benjaminpeterson/Documents/research/BLiMMP/results/incubations/methylation/incubations2019/BLiMMP_006_per_Me198Hg.png"
include_graphics(graph.path)
```


Let's also look at it as a change in Me^198^Hg over the experiment.
One of our algal-amended samples is pretty high in t=0 sample, so looking at it this way might clean it up.

``` {r change_in_MeHg_BLiMMP_trip_006, fig.width = 10, fig.height = 4, echo=FALSE, message = FALSE}
library(patchwork)

# Generate needed color vector
color.vector.006 <- c(cb.translator["black"],
                      cb.translator["bluishgreen"],
                      cb.translator["vermillion"],
                      cb.translator["orange"],
                      cb.translator["blue"],
                      cb.translator["reddishpurple"])
names(color.vector.006) <- c("unfiltered\nmolybdate",
                             "unfiltered\nunamended",
                             "filtered\nunamended",
                             "unfiltered\nstarch",
                             "unfiltered\nstarch\nmolybdate",
                             "unfiltered\nalgal")


# Read in the data
data.path <- "/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/incubations/MeHg/incubations2019_Me198Hg_production.csv"
MeHg.data <- read.csv(data.path,
                      stringsAsFactors = FALSE) %>%
  filter(tripID == "BLiMMP_trip_006") %>%
  select(incubationID, depth, treatment, Me198Hg_production)
plot.list <- list()
for (depth.of.interest in unique(MeHg.data$depth)) {
  depth.of.interest.m <- paste(depth.of.interest, "m", sep = "")
  plot.list[[depth.of.interest.m]] <- MeHg.data %>%
    filter(depth == depth.of.interest) %>%
    mutate(treatment = gsub("-", "\n", treatment)) %>%
    mutate(treatment = factor(treatment, levels =  c("filtered\nunamended",
                                                     "unfiltered\nunamended",
                                                     "unfiltered\nmolybdate",
                                                     "unfiltered\nalgal",
                                                     "unfiltered\nstarch",
                                                     "unfiltered\nstarch\nmolybdate"))) %>%
    ggplot(aes(x = treatment,
               y = Me198Hg_production,
               color = treatment)) +
    geom_jitter(position = position_jitter(0.1)) +
    scale_color_manual(values = color.vector.006) +
    theme_classic() +
    theme(legend.position = "none") +
    ylim(min(MeHg.data$Me198Hg_production)-0.001,
         max(MeHg.data$Me198Hg_production)+0.003) +
    xlab("") +
    ggtitle(depth.of.interest.m)
    
}

plot.list[[1]] + plot.list[[2]]

```

I also plotted it, again, as the change in percent MeHg:

```{r change_in_fraction_MeHg_BLiMMP_trip_006, echo = FALSE, out.height= "75%", out.width="75%"}

# Read in a PDF of the document
graph.path <- "/Users/benjaminpeterson/Documents/research/BLiMMP/results/incubations/methylation/incubations2019/BLiMMP_006_change_in_per_Me198Hg.png"
include_graphics(graph.path)
```


Some of the findings from our previous incubations are confirmed here.
First, there is no abiotic MeHg production, but we do see MeHg production at both depths with sulfide accumulations.
It seems to be higher in the metalimnion, but we only have two replicates and the data is rather noisy.
We didn't have any anoxic depths without sulfide accumulation to compare, as we did at the other dates.
It's impossible to say for sure with the small number of samples, but it does look like MeHg production might be a little higher in the upper sample (16.2m), which is consistent with the higher heterotrophic activity at that depth.
Once again, molybdate addition completely eliminates MeHg production.

Surprisingly, the starch and algal amendments result in a decrease in MeHg production.
The algal biomass resulted in a greater decrease of methylation than starch at 16.2m.
At 19.3m, one of the starch-amended incubations had lower MeHg production, but we also saw a complete loss of sulfide in that bag (see below).
Finally, adding starch and molybdate also completely knocked out methylation.
This response is curious.

One potential explanation would be if the injection resulted in a change in the redox status of the bags.
We didn't inject sterile water into the ambient conditions as a way to control for effects of just injecting anything into the sample, likely as a way for O~2~ to get in, but it is unlikely that this would have that large of an impact.
Besides, we injected the Hg tracers in using the same method.
But, just to be sure, I checked the sulfide levels in the bags, below.

```{r sulfide_data_incubations_BLiMMP_trip_006,  echo=FALSE, message = FALSE, fig.height = 5, fig.width = 10}

trip.of.interest <- "BLiMMP_trip_006"
# Generate needed color vector
color.vector.006 <- c(cb.translator["black"],
                      cb.translator["bluishgreen"],
                      cb.translator["vermillion"],
                      cb.translator["orange"],
                      cb.translator["blue"],
                      cb.translator["reddishpurple"])
names(color.vector.006) <- c("unfiltered-molybdate",
                             "unfiltered-unamended",
                             "filtered-unamended",
                             "unfiltered-starch",
                             "unfiltered-starch-molybdate",
                             "unfiltered-algal")

par(mfrow = c(1, 2),
    mar = c(4.5, 4.5, 3, 1))

sulfide.data <- read.csv("/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/waterChemistry/sulfide/MA_data.csv",
                         stringsAsFactors = FALSE) %>%
  filter(tripID == trip.of.interest,
         depth == 16.18) %>%
  left_join(MeHg.data) %>%
  select(incubationID, treatment, S_conc_uM, Me198Hg_production)

WC.S.levels <- read.csv("/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/waterChemistry/sulfide/WC_data.csv",
                                   stringsAsFactors = FALSE) %>%
            filter(tripID == trip.of.interest,
                   depth == 16.18) %>%
            select(S_conc_uM) %>%
            unlist(use.names = FALSE)
plot(x = jitter(sulfide.data$S_conc_uM),
     y = jitter(sulfide.data$Me198Hg_production),
     col = color.vector.006[sulfide.data$treatment],
     pch = 18,
     ylim = c(-0.01, max(sulfide.data$Me198Hg_production)*1.2),
               xlim = c(0, max(sulfide.data$S_conc_uM)*1.2),
               xlab = "Sulfide concentration\nat end of incubation(uM)",
               ylab = "MeHg production after 24hr (ng/L/day)",
               main = paste("MeHg production compared\nto sulfide levels at 16.2m\nWater column = ",
                            WC.S.levels,
                            "µM",
                            sep = ""),
               cex.main = 0.85)

# Second depth
sulfide.data <- read.csv("/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/waterChemistry/sulfide/MA_data.csv",
                         stringsAsFactors = FALSE) %>%
  filter(tripID == trip.of.interest,
         depth == 19.3) %>%
  left_join(MeHg.data) %>%
  select(incubationID, treatment, S_conc_uM, Me198Hg_production)

WC.S.levels <- read.csv("/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/waterChemistry/sulfide/WC_data.csv",
                                   stringsAsFactors = FALSE) %>%
            filter(tripID == trip.of.interest,
                   depth == 19.3) %>%
            select(S_conc_uM) %>%
            unlist(use.names = FALSE)
anomalous.point <- sulfide.data %>%
  filter(incubationID == "BLiMMP_MA_0076")
plot(x = jitter(sulfide.data$S_conc_uM),
     y = jitter(sulfide.data$Me198Hg_production),
     col = color.vector.006[sulfide.data$treatment],
     pch = 18,
     ylim = c(-0.01, max(sulfide.data$Me198Hg_production)*1.2),
               xlim = c(0, max(sulfide.data$S_conc_uM)*1.2),
               xlab = "Sulfide concentration\nat end of incubation(uM)",
               ylab = "MeHg production after 24hr (ng/L/day)",
               main = paste("MeHg production compared\nto sulfide levels at 19.3m\nWater column = ",
                            WC.S.levels,
                            "µM",
                            sep = ""),
               cex.main = 0.85)
text(x = anomalous.point$S_conc_uM + 15,
     y = anomalous.point$Me198Hg_production + 0.004,
    labels = anomalous.point$incubationID,
    col = color.vector.006[anomalous.point$treatment])

```

There is no definitive trend in the sulfide data.
At 16.2m, it might be the case within conditions that lower sulfide leads to less MeHg production, but it's impossible to say with any certainty with the data we have, with limited sample number.
At 19.3m, there is definitely no trend between lost sulfide and MeHg production.
The one sample that we did see tank on MeHg production unexpectedly was BLiMMP_MA_0076, which was a starch amended incubation.
That incubation was also completely out of sulfide, suggesting some O~2~ intrusion which would explain the lack of methylation.

I had hypothesized that the amendments of starch and algal biomass would have amped up MeHg production due to a ramping up of whole community metabolism, so let's check out the leucine uptake data.
We'll look at this as µg of bacterial carbon production per liter per hour, which we've calculated based on the leucine uptake rates, same as we did for the water column samples.
We only did this from incubations at 16.18m on 2019-10-08 (BLiMMP_sample_0034).
Let's take the average of the original leucine uptake assay and get an average for it, then find the fractional change between that and each of our samples, just to see how different they really are.
We took two samples from each incubation, so two dots correspond to one bag.

```{r Shift_in_leucine_uptake, width = 10, height = 10, echo=FALSE}

original.uptake <- read.csv("/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/leucineUptake/water_column_uptake.csv",
                            stringsAsFactors = FALSE) %>%
  filter(sampleID == "BLiMMP_sample_0034") %>%
  select(µgBCP_per_L_hr) %>%
  unlist(use.names = FALSE) %>%
  mean()

leucine.uptake.MA <- read.csv("/Users/benjaminpeterson/Documents/research/BLiMMP/dataEdited/leucineUptake/incubation_uptake.csv",
                              stringsAsFactors = FALSE) %>%
  select(incubationID, treatment, µgBCP_per_L_hr) %>%
  mutate(delta_µgBCP_per_L_hr =( (µgBCP_per_L_hr - original.uptake) / original.uptake)*100) %>%
  filter(treatment != "unamended - filtered")

# stripplot by group

# Generate needed color vector
color.vector.006 <- c(cb.translator["black"],
                      cb.translator["bluishgreen"],
                      cb.translator["vermillion"],
                      cb.translator["orange"],
                      cb.translator["blue"],
                      cb.translator["reddishpurple"])
names(color.vector.006) <- c("unfiltered\nmolybdate",
                             "unfiltered\nunamended",
                             "filtered\nunamended",
                             "unfiltered\nstarch",
                             "unfiltered\nstarch\nmolybdate",
                             "unfiltered\nalgal")

# Generate plot
leucine.uptake.MA %>%
  mutate(treatment = gsub("-", "\n", treatment)) %>%
  mutate(treatment = factor(treatment, levels =  c("filtered\nunamended",
                                                   "unfiltered\nunamended",
                                                   "unfiltered\nmolybdate",
                                                   "unfiltered\nalgal",
                                                   "unfiltered\nstarch",
                                                   "unfiltered\nstarch\nmolybdate"))) %>%
  ggplot(aes(x = treatment,
             y = delta_µgBCP_per_L_hr,
             color = treatment)) +
  geom_jitter(position = position_jitter(0.1)) +
  scale_color_manual(values = color.vector.006) +
  theme_classic() +
  theme(legend.position = "none") +
  ylim(-100, 25) +
  xlab("") +
  ylab("Change in µgBCP / L / hr")
    

```

Not super informative, unfortunately.
BCP was almost completely knocked out in the filtered controls, as expected.
In the other three conditions that we did this for, the data is pretty noisey.
But, it seems that there is a general drop-off in BCP over the course of the incubation, anywhere from 0-40%.
This drop-off was very consistent between the ambient conditions and the starch amendment.
If anything, the molybdate-amended sample retained more of the ambient heterotrophy, but again, the data is pretty noisey.
If this analysis is run again, I think the samples need to be sure to be processed ASAP, since it's tough to rule out effects of delayed processing on this (incubation samples were processed the next afternoon).




## Takeaways:

1. Boom, again we got detectable methylation. 
2. MeHg production seems to be slightly higher near the oxic/anoxic interface, tracking with heterotrophic activity, but data is pretty noisy. More replicates next time.
3. Molybdate once again knocks out MeHg production.
4. Starch and algal amendments *decrease* MeHg production. While we only have two reps from each condition at each depth, this seems to be a consistent result. The decrease isn't explained by sulfide levels or changes in heterotrophy.
5. Gonna need more replicates overall. Saying anything about different MeHg production at different depths is gonna need more reps. In the future, split up depth discrete sampling from testing different amendments, to be able to give them each the replicates they deserve.
