---
title: "Demethylation analyses"
author: "Benjamin D. Peterson"
date: "2024-06-17"
output: html_document
---

```{r setup, include = FALSE}
rm(list = ls())
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(echo = TRUE)
new.wd <- gsub("\\/[[:alpha:]]+$", "", getwd())
knitr::opts_knit$set(root.dir = new.wd)
rm(new.wd)
```

```{r read_in_color_translator, include = FALSE}
cb.translator <- readRDS("references/colorblind_friendly_colors.rds")
```

# Prepare data

First we'll read in the merB data:

```{r read_in_merB_data}
merB_data <- read.csv("dataFinal/merB_data.csv")
```

Then we'll read in the metadata and generate a useful table to link to the abundance data.

```{r prepare_MG_MT_metadata}
MG_metadata <- read.csv("metadata/metagenome_metadata.csv") %>%
  dplyr::rename(omicID = metagenomeID) %>%
  mutate(seqType = "MG") %>%
  select(omicID, sampleID, startDate, depth, seqType)
MT_metadata <- read.csv("metadata/metatranscriptome_metadata.csv") %>%
  dplyr::rename(omicID = metatranscriptomeID) %>%
  mutate(seqType = "MT") %>%
  select(omicID, sampleID, startDate, depth, seqType)
omic_metadata <- rbind(MG_metadata,
                       MT_metadata) %>%
  mutate(date_depth = paste(startDate, ":", depth, "m",
                            sep = "")) %>%
  select(omicID, date_depth)
rm(MG_metadata, MT_metadata)
```

We'll also want to pull in the sulfide data.

```{r prepare_sulfide_data, message = FALSE}
# Read in sulfide data
sulfide_data <- read.csv("dataFinal/water_chem_data.csv") %>%
  group_by(date, depth) %>%
  summarize(sulfide_uM = mean(sulfide_uM, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(date_depth = paste(date, ":", depth, "m",
                            sep = "")) %>%
  select(date_depth, sulfide_uM)
```

Finally, we'll read in the K~dem~ data.

```{r prepare_kmet_incubation_data, message = FALSE}
# Read in Kmet data
Kdem_data <- read.csv("dataFinal/incubation_Hg_rate_data.csv") %>%
  mutate(date_depth = paste(startDate, ":", depth, "m",
                            sep = "")) %>%
  filter(date_depth != "2020-10-10:15.7m") %>%
  select(date_depth, treatment, Kdem_slope)

Kdem_data_summarized <- Kdem_data %>%
  group_by(date_depth, treatment) %>%
  summarize(Kdem_slope_mean = mean(Kdem_slope, na.rm = TRUE),
            Kdem_slope_sd = sd(Kdem_slope, na.rm = TRUE),
            Kdem_count = n(),
            Kdem_slope_sem = Kdem_slope_sd / sqrt(Kdem_count))
```


# Check demethylation rates

Before looking at the *merB* genes, we'll want to check out the demethylation rates. 

```{r check_kmet_incubation_rates, message = FALSE}
naming_vector_kdem <- c("Ambient", "Molybdate-treated", "Filtered control")
names(naming_vector_kdem) <- c("unfiltered-unamended", "unfiltered-molybdate", "filtered-unamended")
color_vector_kdem <- cb.translator[c("bluishgreen", "black", "orange")]
names(color_vector_kdem) <- names(naming_vector_kdem)

date_depth_order_by_sulfide <- sulfide_data %>%
  filter(date_depth %in% Kdem_data_summarized$date_depth) %>%
  arrange(sulfide_uM) %>%
  select(date_depth) %>%
  unlist()
names(date_depth_order_by_sulfide) <- NULL

Kdem_data_summarized %>%
  mutate(treatment = fct_relevel(treatment, names(naming_vector_kdem)),
         date_depth = as.factor(date_depth),
         date_depth = fct_relevel(date_depth, date_depth_order_by_sulfide)) %>%
  ggplot(aes(x = date_depth,
             y = Kdem_slope_mean,
             fill = treatment)) + 
  geom_bar(stat = "identity",
           color = "black", 
           position = position_dodge()) +
  scale_fill_manual(values = color_vector_kdem,
                    labels = naming_vector_kdem) +
  geom_errorbar(aes(ymin = Kdem_slope_mean - Kdem_slope_sem,
                    ymax = Kdem_slope_mean + Kdem_slope_sem),
                width = 0.2,
                position = position_dodge(0.9)) +
  labs(x = element_blank(),
       y = expression(Kdem)) +
  theme_bw() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
```

This is extremely confusing data.
First of all, the filtered controls showed high Kdem, especially at higher sulfide concentrations.
In 2021, the Kdem values under ambient conditions were comparable to the filtered control conditions.
However, in 2020, there was nearly no demethylation occurring under ambient conditions regardless of sulfide.
To throw the cherry on top, the addition of molybdate stopped MeHg degradation regardless of the year or sulfide concentration.
This indicates multiple competing processes.
Photodemethylation can be ruled out, and the *merB* gene is unlikely to be very active.
Hg(0) concentrations are low in the anoxic hypolimnion, suggesting that reductive demethylation is unlikely to occur.
This leaves oxidative demethylation.
Some abiotic or 

# Plot *merB* data against environmental variables

Set up the merB data generally:

```{r prepare_merB_data, message = FALSE }
merB_plotting_data <- merB_data %>%
  filter(clstr_rep == 1) %>%
  select(clstr_rep, divergent_seq,
         all_of(grep("_MG_", names(merB_data), value = TRUE)),
         all_of(grep("_MT_", names(merB_data), value = TRUE))) %>%
  gather(key = omicID,
         value = coverage,
         c(all_of(grep("_MG_", names(merB_data), value = TRUE)),
         all_of(grep("_MT_", names(merB_data), value = TRUE)))) %>%
  group_by(divergent_seq, omicID) %>%
  summarise(coverage = sum(coverage)) %>%
  ungroup()
```


### *merB* vs. sulfide

First check the *merB* abundance and transcription against sulfide concentrations.
Plot both the overall *merB* coverages (in gray) and the divergent sequences (red) and non-divergent sequences (blue) separately (still on the same plot though).

```{r plot_merB_vs_sulfide, echo = FALSE, message = FALSE, fig.width = 7.5, fig.height = 5 }
year_vector <- c(21, 24)
names(year_vector) <- c("2020", "2021")

color_vector <- c("red", "blue")
names(color_vector) <- c("TRUE", "FALSE")

sulfide_vs_merB <- function(omic_type,
                               ylab_to_use) {
  plot_data <- merB_plotting_data %>%
    filter(grepl(omic_type, omicID)) %>%
    group_by(omicID) %>%
    summarise(coverage = sum(coverage)) %>%
    left_join(omic_metadata) %>%
    left_join(sulfide_data) %>%
    filter(date_depth != "2020-10-10:15.7m")
  plot(x = plot_data$sulfide_uM,
       y = plot_data$coverage,
       pch = year_vector[as.character(year(strsplit(plot_data$date_depth, ":") %>%
                                             sapply("[", 1)))],
       col = "gray25",
       bg = "gray85",
       lwd = 1.5,
       cex = 2,
       xlab = "Sulfide (µM)",
       ylab = ylab_to_use)
  
  plot_data <- merB_plotting_data %>%
    filter(grepl(omic_type, omicID)) %>%
    left_join(omic_metadata) %>%
    left_join(sulfide_data)
  points(x = plot_data$sulfide_uM,
         y = plot_data$coverage,
         pch = 16,
         col = color_vector[as.character(plot_data$divergent_seq)])
}

#### Set up plots ####
par(mfrow = c(1, 2),
    mar = c(3, 3, 1, 1),
    tck = -0.008,
    mgp = c(1.5, 0.2, 0))
sulfide_vs_merB(omic_type = "MG",
                ylab_to_use = "merB abundance (%)")
sulfide_vs_merB(omic_type = "MT",
                ylab_to_use = "merB transcription (10^6 copies/L)")
legend("topleft",
       legend = names(year_vector),
       pch = year_vector)

```

We do observe a general increase in both *merB* abundance and transcription with sulfide concentration, which is interesting. This is nearly entirely driven by these divergent sequences, however, especially the gene abundance.
The non-divergent sequences do show a bump in transcription near the middle of the sulfide gradient.

### *merB* vs. Kdem

```{r plot_merB_vs_Kdem, echo = FALSE, message = FALSE, fig.width = 7.5, fig.height = 4 }
year_vector <- c(21, 24)
names(year_vector) <- c("2020", "2021")

color_vector <- c("red", "blue")
names(color_vector) <- c("TRUE", "FALSE")

merB_vs_Kdem <- function(omic_type,
                         ylab_to_use,
                         ylim_to_use = NULL) {
  plot_data <- merB_plotting_data %>%
    filter(grepl(omic_type, omicID)) %>%
    group_by(omicID) %>%
    summarise(coverage = sum(coverage, na.rm = TRUE)) %>%
    left_join(omic_metadata) %>%
    group_by(date_depth) %>%
    summarise(coverage_mean = sum(coverage, na.rm = TRUE),
              coverage_sd = sd(coverage, na.rm = TRUE)) %>%
    left_join(Kdem_data) %>%
    filter(date_depth != "2020-10-10:15.7m")
  plot(x = plot_data$Kdem_slope_mean,
       y = plot_data$coverage_mean,
       cex = 0,
       xlim = c(-0.1, 0.3),
       ylim = ylim_to_use,
       xlab = "Kdem (day^-1)",
       ylab = ylab_to_use)
  arrows(x0 = plot_data$Kdem_slope_mean - plot_data$Kdem_slope_sd,
         x1 = plot_data$Kdem_slope_mean + plot_data$Kdem_slope_sd,
         y0 = plot_data$coverage_mean, y1 = plot_data$coverage_mean,
         code = 3, lwd = 1, col = "black", angle = 90,
         length = 0.05)
  arrows(x0 = plot_data$Kdem_slope_mean, x1 = plot_data$Kdem_slope_mean,
         y0 = plot_data$coverage_mean - plot_data$coverage_sd,
         y1 = plot_data$coverage_mean + plot_data$coverage_sd,
         code = 3, lwd = 1, col = "black", angle = 90,
         length = 0.05)
  points(x = plot_data$Kdem_slope_mean,
         y = plot_data$coverage_mean,
         pch = year_vector[as.character(year(strsplit(plot_data$date_depth, ":") %>%
                                               sapply("[", 1)))],
         col = "gray25",
         bg = "gray85",
         ylim = ylim_to_use,
         lwd = 1.5,
         cex = 2,
         )
  }

#### Set up plots ####
par(mfrow = c(1, 2),
    mar = c(3, 3, 1, 1),
    tck = -0.008,
    mgp = c(1.5, 0.2, 0))
merB_vs_Kdem(omic_type = "MG",
             ylab_to_use = "merB abundance (%)",
             ylim_to_use = c(0, 7.5))
merB_vs_Kdem(omic_type = "MT",
                ylab_to_use = "merB transcription (10^6 copies/L)",
             ylim_to_use = c(0, 8))
legend(x = -0.3, y = 7,
       legend = names(year_vector),
       pch = year_vector,
       col = "gray25",
       pt.bg = "gray85")
```

Well, there is clearly more *merB* in the regions with more demethylation occurring.
Interesting, there's only demethylation really occurring in 2021.