---
title: "hgcA transcription analysis"
author: "Benjamin D. Peterson"
date: "2023-11-17"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
#setwd("/Users/benjaminpeterson/Documents/research/BLiMMP")
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(echo = TRUE)
new.wd <- gsub("\\/[[:alpha:]]+$", "", getwd())
knitr::opts_knit$set(root.dir = new.wd)
rm(new.wd)
```

```{r color_vector, include=FALSE}
cb_translator <- readRDS("references/colorblind_friendly_colors.rds")
```

# Set up data sheets

First, pull in the metatranscriptomics metadata.

```{r read_in_metadata}
MT_metadata <- read.csv("metadata/metatranscriptome_metadata.csv") %>%
  dplyr::rename(mt_id = metatranscriptomeID) %>%
  mutate(seqType = "MT") %>%
  mutate(date_depth = paste(startDate, ":", depth, "m",
                            sep = "")) %>%
  select(mt_id, date_depth)

```

Then, pull in the *gyrB* gene-to-bin information and metatranscriptome data for all genes.
Combine that to get *gyrB* info per genome.

```{r reading_in_gyr_data, warning=FALSE, message=FALSE}
gyrB_bin_info <- read.table("dataEdited/bin_based_analyses/final_bin_data/gyrB_to_bin.tsv",
                            sep = '\t', col.names = c("gene_id", "bin_id"))
MT_data <- read.table("dataEdited/bin_based_analyses/final_bin_data/bin_MT_data.tsv",
                      sep = '\t', header = TRUE)
gyrB_bin_data <- left_join(gyrB_bin_info, MT_data)
rm(gyrB_bin_info)
```

Some of the bins have multiple *gyrB* annotations.

```{r investigate_dups_in_gyrB}
dups <- gyrB_bin_data %>%
  filter(duplicated(bin_id)) %>% select(bin_id) %>% unlist(use.names = FALSE)
investigate_dups <- gyrB_bin_data %>% filter(bin_id %in% dups)
investigate_dups
rm(dups)
```

Looks like the dups usually have one that is much more abundant.
The low abundance one is often non-existent.
We'll use the highest coverage of any of them for this analysis.

```{r clean_up_dups_in_gyrB, message = FALSE}
gyrB_ref_data <- gyrB_bin_data %>%
  gather(key = mt_id,
         value = gyr_cov,
         -c(1,2)) %>%
  group_by(bin_id, mt_id) %>%
  summarise(gyr_cov = max(gyr_cov))
```

Now get the hgcA bin data together, joining the *hgcA* gene ID to the bin ID, taxonomy, and metabolic classification. 

```{r hgcA_data_prep, message = FALSE}
hgcA_bin_info <- read.table("dataEdited/bin_based_analyses/final_bin_data/hgcA_to_bin.tsv",
                            sep = '\t', col.names = c("gene_id", "bin_id"))
bin_data <- readRDS(file = 'dataEdited/bin_based_analyses/bin_data_aggregate_2_metabolism.rds') %>%
  mutate(phylum = gtdb_tax %>%
           strsplit("p__") %>% sapply('[', 2) %>%
           strsplit(";c__") %>% sapply('[', 1)) %>%
  select(bin_id, cluster_ID, hms_id, phylum, meta_code) %>%
  left_join(hgcA_bin_info)
rm(hgcA_bin_info)
# Two of these hgcA sequences are in the same bin, need to parse those out:
bin_data[which(bin_data$gene_id == "BLI21_assembly105_000000000783_11"), "cluster_ID"] <- "BLI_hgcA_clstr_0056"
bin_data[which(bin_data$gene_id == "BLI21_assembly105_000000003111_6"), "cluster_ID"] <- "BLI_hgcA_clstr_0066"
```

Next we'll add the transcriptional regulation information generated in the `hgcA_GN.Rmd` document.

```{r add_transcriptional_regulation_info}
TR_data <- readRDS("working_directory/transcriptional_regulation_vector.rds")
bin_data <- bin_data %>%
  mutate(TR_assignment = TR_data[cluster_ID])
attr(bin_data$TR_assignment, "names") <- NULL
```

Next we'll read in the sulfide data.

```{r read_in_sulfide_data}
#### Read in geochem data ####
sulfide_data <- read.csv("dataFinal/water_chem_data.csv") %>%
  mutate(date_depth = paste(date, ":", depth, "m",
                            sep = "")) %>%
  group_by(date_depth) %>%
  summarize(sulfide_uM = mean(sulfide_uM, na.rm = TRUE)) %>%
  ungroup()
```

Finally we'll combine all the data together.

```{r combine_all_data, message = FALSE }
hgcA_bin_data <- left_join(bin_data %>%
                             select(bin_id, gene_id, hms_id, phylum, meta_code, TR_assignment),
                           MT_data %>%
                             filter(gene_id %in% bin_data$gene_id)) %>%
  gather(key = mt_id,
         value = hgcA_cov,
         -c(1:6)) %>%
  left_join(gyrB_ref_data) %>%
  filter(!is.na(gyr_cov),
         gyr_cov != 0) %>%
  mutate(hgcA_gyrB_ratio = hgcA_cov / gyr_cov) %>%
  left_join(MT_metadata) %>%
  # group_by(bin_id, date_depth, hgcA_gyrB_ratio) %>%
  # left_join(hgcA_taxonomy) %>%
  left_join(sulfide_data)
```

One of these bins has two *hgcA* sequences, let's see how they compare:

```{r compare_two_genes_same_bin}
hgcA_bin_data %>%
  filter(gene_id %in% c("BLI21_assembly105_000000000783_11", "BLI21_assembly105_000000003111_6"),
         hgcA_gyrB_ratio > 0) %>%
  group_by(date_depth, gene_id) %>%
  summarise(count = n())

hgcA_bin_data %>%
  filter(gene_id %in% c("BLI21_assembly105_000000000783_11", "BLI21_assembly105_000000003111_6")) %>%
  ggplot(aes(x = sulfide_uM,
             y = hgcA_gyrB_ratio,
             group = gene_id,
             color = gene_id,
             shape = gene_id)) +
  geom_point() +
  xlim(c(0, 150)) +
  ylab("hgcA:gyrB transcription") +
  xlab("Sulfide (µM)") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```
This data is super noisy actually, very few locations where the *hgcA* gene has any transcripts.
Overall, the 783_11 gene is more transcribed, but it's noisy.
Neither has a preceding transcriptional regulator, so no need to consider that.
We'll just keep both of them in the dataset, as it shouldn't change much regarding how the downstream analyses are done.


# Dealing with 0's. 

Okay, in later analyses I was running into some serious issues with 0's.
The 0 ratios are throwing off the analyses, and since the *hgcA* gene seems to be on the lower end of transcription, we're getting several genes where there is *gyr* but no *hgcA*.
We can't just remove the zeros, since that will skew the analysis; rather, I need to filter out the lower abundance genes, or more specifically, the genomes that are low enough in abundance to be susceptible to having zeros.
How to identify the threshold though?
What if we take the abundance of each hgcA+ genome and plot it on the x-axis, while on the y-axis we plot the number of 0 values in the triplicate metatranscriptomes?
To do this, first we'll need to calculate the genome abundance at each date-depth.
Actually, we didn't calculate the genomes depths, so instead I'll use the *hgcA* gene.

```{r genome_abundance_at each date-depth, message = FALSE}
MG_data <- read.csv("dataFinal/hgcA_data.csv") 
MG_data <- MG_data %>%
  select(seqID, all_of(grep("_MG_", names(MG_data),
                            value = TRUE))) %>%
  gather(key = metagenomeID,
         value = MG_coverage,
         -1) %>%
  left_join(read.csv("metadata/metagenome_metadata.csv") %>%
              mutate(date_depth = paste(startDate, ":", depth, "m",
                                        sep = "")) %>%
              select(metagenomeID, date_depth)) %>%
  mutate(scaffold_id = paste(seqID %>% strsplit("_") %>% sapply("[", 1),
                             seqID %>% strsplit("_") %>% sapply("[", 2),
                             seqID %>% strsplit("_") %>% sapply("[", 3),
                             sep = "_")) %>%
  group_by(scaffold_id, date_depth) %>%
  summarise(MG_coverage = mean(MG_coverage)) %>%
  ungroup()
```

Next, look at the number of MT replicates at each date-depth that have *hgcA* hits (hgcA:gyrA is above 0) and plot this against the gene abundance.

```{r zero_counts, message = FALSE}
zero_counts_data <- hgcA_bin_data %>%
  filter(hgcA_gyrB_ratio > 0) %>%
  group_by(date_depth, gene_id, bin_id) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  # group_by(date_depth, gene_id) %>%
  # summarise(count = as.integer(mean(count))) %>%
  # ungroup() %>%
  mutate(scaffold_id = paste(gene_id %>% strsplit("_") %>% sapply("[", 1),
                             gene_id %>% strsplit("_") %>% sapply("[", 2),
                             gene_id %>% strsplit("_") %>% sapply("[", 3),
                             sep = "_"))
zero_counts_data <- zero_counts_data %>%
  inner_join(MG_data,
            by = c("date_depth", "scaffold_id")) %>%
  select(-scaffold_id)
zero_counts_data %>%
  ggplot(aes(x = MG_coverage,
             y = count)) +
  geom_jitter(height = 0.1) +
  theme_bw() +
  scale_x_log10()
```

Hmm, this did not show a clear cutoff point.
Let's see how many genomes we have with hits in all three MTs at multiple sites.

```{r bins_with_hgcA_hits_all_MTs }
zero_counts_data %>%
  filter(count == 3) %>%
  group_by(gene_id, bin_id) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  # Keep ones that appear in all three replicates at least 4 times
  filter(count >= 2) %>%
  left_join(bin_data)
  
```

Only one HMS (BLI_hgcA_HMS_0012) has hits in all metatranscriptomes we sequenced, with the exception of the one with no *hgcA*.
This HMS is a SRB and has arsR.
BLI_hgcA_HMS_0019 has hits in all three metatranscriptomes replicates from 5 locations, while BLI_hgcA_HMS_0009 has hits from all three metatranscriptomes replicates from 4 locations.
Let's look more closely at these before diving into the others.

```{r BLI_hgcA_HMS_0012_TR, message = FALSE }
HMS_TO_PLOT <- "BLI_hgcA_HMS_0012"
plot_HMS_hgcA_rel_expr_by_sulfide <- function(HMS_TO_PLOT) {
  HMS_info <- hgcA_bin_data %>%
    filter(hms_id == HMS_TO_PLOT,
           hgcA_gyrB_ratio > 0) %>%
    group_by(hms_id) %>%
    summarise(phylum = unique(phylum),
              TR_assignment = unique(TR_assignment)) %>%
    ungroup() %>%
    select(-hms_id) %>%
    unlist(use.names = FALSE) %>%
    paste(., collapse = ":")
  hgcA_bin_data %>%
    filter(hms_id == HMS_TO_PLOT,
           hgcA_gyrB_ratio > 0) %>%
    group_by(bin_id, sulfide_uM) %>%
    summarise(hgcA_gyrB_ratio_mean = mean(hgcA_gyrB_ratio),
              hgcA_gyrB_ratio_sd = sd(hgcA_gyrB_ratio),
              hgcA_gyrB_ratio_count = n(),
              hgcA_gyrB_ratio_sem = hgcA_gyrB_ratio_sd / sqrt(hgcA_gyrB_ratio_count)) %>%
    ggplot(aes(x = sulfide_uM,
               y = hgcA_gyrB_ratio_mean,
               group = bin_id,
               color = bin_id)) +
    geom_point() +
    geom_line() +
    geom_errorbar(aes(ymin = hgcA_gyrB_ratio_mean - hgcA_gyrB_ratio_sem,
                      ymax = hgcA_gyrB_ratio_mean + hgcA_gyrB_ratio_sem),
                  width = 2) +
    xlim(c(0, 150)) +
    ylab("hgcA:gyrB transcription") +
    theme_bw() + 
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
    scale_y_log10(limits = c(0.01, 10)) +
    theme(legend.position = "none") +
    labs(title = HMS_info)
}
ggarrange(plot_HMS_hgcA_rel_expr_by_sulfide("BLI_hgcA_HMS_0012"),
          plot_HMS_hgcA_rel_expr_by_sulfide("BLI_hgcA_HMS_0019"),
          plot_HMS_hgcA_rel_expr_by_sulfide("BLI_hgcA_HMS_0009"),
          plot_HMS_hgcA_rel_expr_by_sulfide("BLI_hgcA_HMS_0016"),
          nrow = 2,
          ncol = 2)
```

Finally, some of these are going to be affected by the fact that there are multiple bins representing one organism.
We'll need to combine this information by HMS. 

# hgcA expression by transcriptional regulation


```{r hgcA_expression_by_transcriptional_regulation, message = FALSE }

hgcA_bin_data %>%
  filter(hgcA_gyrB_ratio > 0) %>%
  group_by(hms_id, mt_id, sulfide_uM, TR_assignment) %>%
  summarize(hgcA_gyrB_ratio = mean(hgcA_gyrB_ratio)) %>%
  ungroup() %>%
  group_by(hms_id, sulfide_uM, TR_assignment) %>%
  summarise(hgcA_gyrB_ratio_mean = mean(hgcA_gyrB_ratio),
            hgcA_gyrB_ratio_sd = sd(hgcA_gyrB_ratio),
            hgcA_gyrB_ratio_count = n(),
            hgcA_gyrB_ratio_sem = hgcA_gyrB_ratio_sd / sqrt(hgcA_gyrB_ratio_count)) %>%
  ggplot(aes(x = sulfide_uM,
             y = hgcA_gyrB_ratio_mean,
             group = hms_id,
             color = TR_assignment)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = hgcA_gyrB_ratio_mean - hgcA_gyrB_ratio_sem,
                    ymax = hgcA_gyrB_ratio_mean + hgcA_gyrB_ratio_sem),
                width = 2) +
  xlim(c(0, 150)) +
  ylab("hgcA:gyrB transcription") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_y_log10(limits = c(0.01, 10)) +
  theme(legend.position = "none") +
  labs(title = HMS_info)


hgcA_bin_data %>%
  filter(hgcA_gyrB_ratio > 0) %>%
  group_by(hms_id, mt_id, sulfide_uM, TR_assignment) %>%
  summarize(hgcA_gyrB_ratio = mean(hgcA_gyrB_ratio)) %>%
  ungroup() %>%
  group_by(hms_id, sulfide_uM, TR_assignment) %>%
  summarise(hgcA_gyrB_ratio_mean = mean(hgcA_gyrB_ratio),
            hgcA_gyrB_ratio_sd = sd(hgcA_gyrB_ratio),
            hgcA_gyrB_ratio_count = n(),
            hgcA_gyrB_ratio_sem = hgcA_gyrB_ratio_sd / sqrt(hgcA_gyrB_ratio_count)) %>%
  ggplot(aes(x = sulfide_uM,
             y = hgcA_gyrB_ratio_mean,
             group = TR_assignment,
             color = TR_assignment)) +
  geom_boxplot() 
```















Next I'm 

```{r plot_by_transcription_reg, message = FALSE, echo = FALSE }
TR_data <- hgcA_bin_data %>%
  filter(TR_assignment != "unknown",
         date_depth != "2020-10-10:15.7m") %>%
  group_by(bin_id, date_depth, TR_assignment, sulfide_uM)%>%
  summarise(hgcA_gyrB_ratio = mean(hgcA_gyrB_ratio)) %>%
  arrange(sulfide_uM, date_depth) %>%
  mutate(date_depth = factor(date_depth, unique(date_depth)))

plot_data <- TR_data
plot_data$hgcA_gyrB_ratio[which(plot_data$hgcA_gyrB_ratio == 0)] <- 0.005
plot_data %>%
  ggplot(aes(x = date_depth,
             y = hgcA_gyrB_ratio,
             color = TR_assignment)) +
  geom_boxplot() +
  # xlim(c(0, 150)) +
  ylab("hgcA:gyrB transcription") +
  xlab("Sulfide (µM)") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_y_log10()


# hgcA_bin_data %>%
#   filter(hgcA_gyrB_ratio > 0) %>%
#   group_by(hms_id, mt_id, date_depth, TR_assignment, sulfide_uM) %>%
#   summarise(hgcA_gyrB_ratio = median(hgcA_gyrB_ratio)) %>%
#   group_by(hms_id, date_depth, TR_assignment, sulfide_uM) %>%
#   summarise(hgcA_gyrB_ratio = mean(hgcA_gyrB_ratio)) %>%
#   ggplot(aes(x = sulfide_uM,
#              y = log(hgcA_gyrB_ratio, 10),
#              group = hms_id,
#              color = TR_assignment)) +
#   geom_point() +
#   geom_line() +
#   # xlim(c(0, 150)) +
#   ylab("hgcA:gyrB transcription") +
#   xlab("Sulfide (µM)") +
#   theme_bw() + 
#   theme(panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank())
```


```{r plot_by_guild }
hgcA_bin_data %>%
  group_by(bin_id, date_depth, sulfide_uM, meta_code) %>%
  summarise(hgcA_gyrB_ratio = mean(hgcA_gyrB_ratio)) %>%
  ggplot(aes(x = as.character(sulfide_uM),
             y = hgcA_gyrB_ratio,
             color = meta_code)) +
  geom_boxplot() +
  # xlim(c(0, 150)) +
  ylab("hgcA:gyrB transcription") +
  xlab("Sulfide (µM)") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```



```{r hgcA_data_plot}
# cairo_pdf("/Users/benjaminpeterson/crashplan/poulinLab/labMeetings/myPresentations/20231114/images/hgcA_transcription_sulfide.pdf",
#           width = 6,
#           height = 4)
color_vector <- cb_translator[c("blue", "black")]
names(color_vector) <- c("SRB", "FERM")
hgcA_bin_data %>%
  filter(sulfide_uM > 25) %>%
  group_by(metabolism, sulfide_uM, date_depth) %>%
  summarise(hgcA_gyrB_ratio = mean(hgcA_gyrB_ratio)) %>%
  ungroup() %>%
  ggplot(aes(x = sulfide_uM,
             y = hgcA_gyrB_ratio,
             group = metabolism,
             color = metabolism)) +
  geom_point() +
  geom_line() +
  ylim(c(0, 0.22)) +
  xlim(c(0, 150)) +
  scale_color_manual(values = color_vector) +
  ylab("hgcA:gyrB transcription") +
  xlab("Sulfide (µM)") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
# hgcA_bin_data %>%
#   filter(sulfide_uM > 25) %>%
#   group_by(metabolism, sulfide_uM, date_depth) %>%
#   summarise(gyr_cov = mean(gyr_cov)) %>%
#   ungroup() %>%
#   ggplot(aes(x = sulfide_uM,
#              y = gyr_cov,
#              group = metabolism,
#              color = metabolism)) +
#   geom_point() +
#   geom_line() +
#   scale_color_manual(values = color_vector) +
#   ylab("gyrB coverage") +
#   xlab("Sulfide (µM)") +
#   theme_bw()
# dev.off()
```







```{r total_transcript_plot}

hgcA_bin_data %>%
  group_by(sulfide_uM, date_depth) %>%
  summarise(hgcA_gyrB_ratio = mean(log(hgcA_gyrB_ratio,10))) %>%
  ungroup() %>%
  ggplot(aes(x = sulfide_uM,
             y = hgcA_gyrB_ratio)) +
  geom_point() +
  geom_line() +
  theme_bw()

```