---
title: "Metabolic assignment of hgcA+ bins"
author: "Benjamin D. Peterson"
date: "2023-12-21"
output: html_document
---

```{r setup, include = FALSE}
rm(list = ls())
#setwd("/Users/benjaminpeterson/Documents/research/BLiMMP")
#new.wd <- "/Users/benjaminpeterson/Documents/research/BLiMMP"
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(echo = TRUE)
new.wd <- gsub("\\/[[:alpha:]]+\\/[[:alpha:]]+$", "", getwd())
knitr::opts_knit$set(root.dir = new.wd)
rm(new.wd)
```

First let's read in the initial bin data, with the completeness, taxonomy, grouping, etc.
This is intended to work on the file from the initial bin data aggregation step.

```{r read_in_hgcA_bin_data }
hgcA_bin_data <- readRDS('dataEdited/bin_based_analyses/bin_data_aggregate_1.rds')
```

### Batch HMM data

First we'll read in the data that we want to use to classify the different bins metabolically.
We'll use two primary data sources for this:
First, we'll use the batch HMM workflow that I've developed.
Second, we'll use KOFAMscan data, using on the confident hits.

```{r read_in_and_prep_data }
batch_hmm_data <- read.table("dataEdited/bin_based_analyses/final_bin_data/batch_HMMs_bin_hits.tsv",
                             sep = '\t',
                             header = TRUE) %>%
  group_by(binID, proteinName) %>%
  summarise(geneIDs = paste0(geneID, collapse = ";")) %>%
  spread(key = proteinName,
         value = geneIDs) %>%
  dplyr::rename(bin_id = binID)
hgcA_bin_metabolism_assignment <- hgcA_bin_data %>%
  left_join(batch_hmm_data)

kofamscan_data <- read.table("dataEdited/bin_based_analyses/final_bin_data/kofam_data.tsv",
                             sep = '\t',
                             header = FALSE,
                             col.names = c("bin_id", "delete1", "gene_id", "KOFAM", "delete2", "delete3", "delete4", "gene_info"))
kofamscan_data <- kofamscan_data %>%
  select(!any_of(grep("delete", names(kofamscan_data),
                   value = TRUE)))
```

# Sulfate-reducing bacteria

First I looked for sulfate-reducing bacteria.

```{r SRBs }
dsr_bin_info <- hgcA_bin_metabolism_assignment %>%
  select(bin_id, hms_id, all_of(grep("dsr", names(hgcA_bin_metabolism_assignment))), "aprA") %>%
  gather(key = gene_name,
         value = name,
         -c(1:2)) %>%
  filter(!is.na(name)) %>%
  spread(key = gene_name,
         value = name)
```

This is pretty clear.
All 11 of these bins, which fit into 2 HMSs, are clear sulfate-reducers, they have *dsrABD*.
All but 2 have *aprA* (and the other bin in the HMS does have it), and a previous screen showed they all had *sat*.
We'll mark all 11 as SRBs.

```{r SRB_bin_list}
binList_srb <- dsr_bin_info$bin_id
```

# Nitrate reducers

Next I looked for nitrate reduction capacity.

```{r nitrate_reducers }
nar_bin_info <- hgcA_bin_metabolism_assignment %>%
  select(bin_id, hms_id, gtdb_tax, all_of(grep("nar", names(hgcA_bin_metabolism_assignment)))) %>%
  gather(key = gene_name,
         value = name,
         -c(1:3)) %>%
  filter(!is.na(name)) %>%
  spread(key = gene_name,
         value = name)
binList_nar <- nar_bin_info$bin_id
```

One Verruco and one Bacteroidetes with respiratory nitrate reduction.
*narG* and *narH* show up in the KOFAMscan data as "nitrate reductase / nitrite oxidoreductase", either the alpha or beta subunit.
Also in the KOFAM scan data is "nitrate reductase (cytochrome)".
None of the genomes has nirS or nirK, so no full denitrification.
However, for both bins, they are one of 3-4 bins in an HMS, but the only bin with the narG:
- For the Bacteroidetes (BLI21_coassembly_anvio_bin_0147), the bin with narG was definitely the most complete.
The contig with narG just has narH on it as well, at least that's annotated by KOFAMscan.
Only other MoOR that was identified in this HMS (BLI_hgcA_HMS_0016) was next two an aconitase subunit.
- For the Verruco (BLI21_coassembly_dasTool_bin_0260), there are 3 other bins in the HMS (BLI_hgcA_HMS_0004). The narG+ one was not manually binned. They all do, however, have cytochrome c oxidase, cytochrome bd oxidase, and complex I, as well as 9-14 MHCs (not many MHCs with 10 multi-heme binding sites, at most 1).

Either way, both of these are probably close enough to be called a high-redox respiratory bin.
However, we'll just stick to calling them RESP, assuming that the bin is capable of respiration and if it's not using sulfate or methanogenesis, it's probably higher up on the redox ladder.

Looked for nitrate reductases in the KOFAM scan results as well.

```{r nitrate_reductase_cytochrome, message = FALSE }
check_data <- kofamscan_data %>%
  filter(grepl("nitrate reductase \\(cytochrome\\)", gene_info)) %>%
  left_join(hgcA_bin_data %>%
              select(bin_id, hms_id, gtdb_tax))
```

Interesting, these are the two sulfate reducers we identified previously.
These are both under the *napA* classification.

# Terminal oxidases

Let's look at what kind of oxidases are present.

Searched for cytochrome o ubiquinol oxidase, called cyo, but there was nothing.

Next checked out the HMMs for cox and cco genes, which both correspond to cytochrome c oxidase, cbb3-type.
They correspond to EC7.1.1.9 (https://www.genome.jp/entry/7.1.1.9), a cytochrome-c oxidase with many different names/subsets.

```{r cco_and_cox_oxidases }
binInfo_cytcOxidase <- hgcA_bin_metabolism_assignment %>%
  select(bin_id, hms_id, gtdb_tax,
         all_of(grep("cco", names(hgcA_bin_metabolism_assignment))),
         all_of(grep("cox", names(hgcA_bin_metabolism_assignment)))) %>%
  gather(key = gene_name,
         value = name,
         -c(1:3)) %>%
  filter(!is.na(name)) %>%
  spread(key = gene_name,
         value = name)
```

The cox HMM picks up on way more genes than the cco one, and all the ones cco picks up on, cox gets as well.
We'll just say that all of these bins have a cytochrome-c oxidase in that class.

```{r cytcOxidase_list }
binList_cytcOxidase <- binInfo_cytcOxidase$bin_id
```

Last I looked at the cydAB genes, correspond to cytochrome bd terminal oxidase.
Got cydA with PF01654.hmm, and cydB with TIGR00203.HMM.

```{r cyd_oxidases }
binInfo_cyd <- hgcA_bin_metabolism_assignment %>%
  select(bin_id, hms_id, gtdb_tax, all_of(grep("cyd", names(hgcA_bin_metabolism_assignment)))) %>%
  gather(key = gene_name,
         value = name,
         -c(1:3)) %>%
  filter(!is.na(name)) %>%
  spread(key = gene_name,
         value = name)
```

Most all of these have both A and B annotated; for the bins with no cydB hit in the batch HMMs, a quick look at kofam suggests that these is a downstream subunit IIl, with one exception.
We'll keep all these bins and describe them as carrying a cytochrome bd terminal oxidase.
This protein has been described as being involved in microaerobic conditions (https://www.ebi.ac.uk/interpro/entry/pfam/PF01654/).

```{r cyd_list }
binList_cyd <- binInfo_cyd$bin_id
```





# Electron transport chain Complex I

I looked for three different types of complex I of the electron transport chain in my bins.
The purpose of looking for these was to identify genomes potentially capable of respiration.
For more information on these complexes, see Kaila and Wikström 2021 (DOI:10.1038/s41579-020-00486-4).

### nuo

First I looked at the batch HMM data for complex I (*nuo*) genes.

```{r nuo }
binInfo_nuo <- hgcA_bin_metabolism_assignment %>%
  select(bin_id, hms_id, gtdb_tax, all_of(grep("nuo", names(hgcA_bin_metabolism_assignment)))) %>%
  gather(key = gene_name,
         value = name,
         -c(1:3)) %>%
  filter(!is.na(name)) %>%
  spread(key = gene_name,
         value = name)
```

However, this data was a little noisy and hard to interpret.
Instead, I looked in the KOFAM scan data and identified bins likely to have complex I.

```{r ETC_manual_ID }

ETC_S1_manual <- c("BLI20_assembly002_dasTool_bin_0004", "BLI20_coassembly_anvio_bin_0009", "BLI20_coassembly_anvio_bin_0027",
                   "BLI20_coassembly_dasTool_bin_0007", "BLI20_coassembly_dasTool_bin_0100", "BLI21_assembly102_anvio_bin_0033",
                   "BLI21_assembly102_dasTool_bin_0076", "BLI21_assembly103_anvio_bin_0017", "BLI21_assembly103_anvio_bin_0023",
                   "BLI21_assembly103_dasTool_bin_0044", "BLI21_assembly103_dasTool_bin_0050", "BLI21_assembly105_anvio_bin_0010",
                   "BLI21_assembly105_dasTool_bin_0059", "BLI21_assembly105_dasTool_bin_0060", "BLI21_assembly105_dasTool_bin_0068",
                   "BLI21_assembly106_anvio_bin_0003", "BLI21_assembly106_anvio_bin_0010", "BLI21_assembly106_dasTool_bin_0028",
                   "BLI21_assembly106_dasTool_bin_0040", "BLI21_assembly106_dasTool_bin_0077", "BLI21_coassembly_anvio_bin_0005",
                   "BLI21_coassembly_anvio_bin_0085", "BLI21_coassembly_anvio_bin_0107", "BLI21_coassembly_anvio_bin_0128",
                   "BLI21_coassembly_anvio_bin_0147", "BLI21_coassembly_dasTool_bin_0029", "BLI21_coassembly_dasTool_bin_0158",
                   "BLI21_coassembly_dasTool_bin_0178", "BLI21_coassembly_dasTool_bin_0253", "BLI21_coassembly_dasTool_bin_0256",
                   "BLI21_coassembly_dasTool_bin_0260", "BLI21_coassembly_dasTool_bin_0275")
```

This was a slog. One thing to note is that many bins have genes annotated as *nuoEF*, but these are not necessarily indicative of a complex I.
A quicker way to do this, and perhaps a gut check, is to select all the "NADH-quinone oxidoreductase" entries, filter out all the *nuoEF* genes, then only keep bins that have more than 3 entries.

```{r ETC_manual_better method}
ETC_S1_auto <- kofamscan_data %>%
  filter(grepl("NADH-quinone oxidoreductase", gene_info)) %>%
  filter(!grepl("subunit [E-F]", gene_info)) %>%
  group_by(bin_id) %>%
  summarise(subunit_count = n()) %>%
  filter(subunit_count > 3) %>%
  select(bin_id) %>%
  unlist(use.names = FALSE)
all(sort(ETC_S1_auto) == sort(ETC_S1_manual))
```

These match up, little extra confidence that the bins were correctly annotated.


### nqr

Next I looked for the nqr gene set, which is the sodium-translocating complex I and contains nqrA, B, C, D, E and F
For this, I just looked in the KOFAMscan data for NADH:ubiquinone oxidoreductase, which I had found to identify Na+-transporting NADH:ubiquinone oxidoreductase subunits.
Many of these bins have just D, E, and/or F subunits, let's look for bins with more than 3 types of subunits.
This pathway shares homology with Rnf.
Nice review by Barquera 2014 on the the biochemistry of this gene.

```{r nqr_search }
nqr_data <- kofamscan_data %>%
  filter(bin_id %in% (kofamscan_data %>%
                        filter(grepl("EC:7.2.1.1", gene_info)) %>% # This returns 204 results, which matches my manual filtering.
                        group_by(bin_id) %>%
                        summarise(nqr_count = n()) %>%
                        filter(nqr_count > 3) %>%
                        select(bin_id) %>%
                        unlist(use.names = FALSE)),
         grepl("EC:7.2.1.1", gene_info)) 
# Summarize what we have
nqr_data_counts <- nqr_data %>%
  group_by(bin_id) %>%
  summarise(nqr_count = n()) %>%
  ungroup()
test <- hgcA_bin_metabolism_assignment %>%
  filter(bin_id %in% nqr_data_counts$bin_id)
```

Decided to label any bin with 4 or more *nqr* subunits as nqr+.
All of the nqr+ bins also have *rnf*.

```{r nqr_designation }
binList_nqr <- nqr_data_counts %>%
  select(bin_id) %>%
  unlist(use.names = FALSE)
```


# Multiheme cytochrome c

One of the metrics I use to determine an obligate fermenter vs. an organism capable of respiration is the presence of multiheme cytochrome c (MHC) genes.
Obligate fermenters tend to have low numbers of MHCs.
MHCs are identified by CXXCH motifs, and I (somewhat arbitrarily) define "multi" as three.
I use a custom python script written by Shaomei He to determine these.
I looked at two metrics: number of MHCs with 3 heme-binding motifs, and numbers of MHCs with 10 heme-binding motifs.

```{r MHC_data, message=FALSE}
MHC_raw_data <- read.table("dataEdited/bin_based_analyses/final_bin_data/heme3_count_bins.tsv",
                       header = TRUE) %>%
  dplyr::rename(bin_id = binID)
MHC_data_3 <- MHC_raw_data %>%
  group_by(bin_id) %>%
  summarize(MHCs_3HBS = n())
MHC_data_10 <- MHC_raw_data %>%
  filter(hemeCount >= 10) %>%
  group_by(bin_id) %>%
  summarize(MHCs_10HBS = n())
MHC_data <- hgcA_bin_data %>%
  select(bin_id, hms_id, gtdb_tax) %>%
  left_join(MHC_data_3,) %>%
  left_join(MHC_data_10) %>%
  mutate(MHCs_3HBS = replace_na(MHCs_3HBS, 0)) %>%
  mutate(MHCs_10HBS = replace_na(MHCs_10HBS, 0))
```

# Metabolic guild assignment


```{r aggregate_data }
assignment_table <- hgcA_bin_data %>%
  select(bin_id, hms_id, hgca_id, completeness, contamination, genome_length, N50, GC,
         gtdb_tax) %>%
  mutate(SRB = (bin_id %in% binList_srb),
         nitrate_reducer = (bin_id %in% binList_nar),
         cyt_c = (bin_id %in% binList_cytcOxidase),
         cyd = (bin_id %in% binList_cyd)) %>%
  left_join(MHC_data)
assignment_table$complexI[(assignment_table$bin_id %in% ETC_S1_manual)] <- "nuo"
assignment_table$complexI[(assignment_table$bin_id %in% binList_nqr)] <- "nqr"
assignment_table$complexI[(assignment_table$bin_id %in% ETC_S1_manual) &
                            (assignment_table$bin_id %in% binList_nqr)] <- "nuo;nqr"
#assignment_table %>%
#  filter(hms_id != "BLI_hgcA_HMS_0006",
#         !(hms_id %in% c("BLI_hgcA_bin_0001", "BLI_hgcA_bin_0004")))
```

There is a strong overlap between bins with high MHC content and the *nuo* complex I.
However, the bins in BLI_hgcA_HMS_0006 have high MHC content (~12-14) and no complex I, either *nuo* or *nqr*.
BLI_hgcA_bin_0001 and BLI_hgcA_bin_0004 also show MHC but no complex I of either type.
They're all in the UBA1067 family, within the RFP12 order in Kiritimatiellae.
Only ones in that family have this scheme.
The bins in HMS BLI_hgcA_HMS_0011 (order UBA8416 in Kiritimatiellae) have no cytochrome c or bd, *nuo*, *dsr*, or *nar*, but do have 4-7 MHCs and an *nqr*.

Conversely, the HMS BLI_hgcA_HMS_0019 (Actinobacteriota, genus Fen713) has three bins that have nuo but few to no MHCs, no cyd, no cytochrome c oxidase.
These three each have a hit against the MoOR HMM, but none of them are assigned in KOFAM, which makes me think they're divergent MoOR genes. 

Otherwise, all bins with two or less MHCs do not have cytochrome c oxidases, cytochrome bd oxidates, complex I machinery, dissimilatory sulfite reductase, or membrane-associated nitrate reductase.
Technically, bin BLI21_coassembly_dasTool_bin_0331 in HMS BLI_hgcA_HMS_0024 does have a gene annotated as cytochrome c oxidase, but the other bin in the HMS does not, and it has no MHCs, so I think the cyt_c annotation is in error.
All of these I will declare as obligately fermentative.

```{r obligate_fermenters}
binAssignment_obligateFermenters <- assignment_table %>%
  filter(MHCs_3HBS <= 2) %>%
  select(bin_id) %>%
  unlist(use.names = FALSE)
```

Now I started to put these together.

```{r add_metabolic_code_to_hgcA_bin_data }
hgcA_bin_data_final <- hgcA_bin_data
hgcA_bin_data_final$meta_code <- NA
hgcA_bin_data_final$meta_code[hgcA_bin_data_final$bin_id %in% binAssignment_obligateFermenters] <- "OB_FERM"
hgcA_bin_data_final$meta_code[hgcA_bin_data_final$bin_id %in% binList_srb] <- "SRB"
hgcA_bin_data_final$meta_code[hgcA_bin_data_final$bin_id %in% binList_nar] <- "FERM"
```

Let's see what we have, and who is lacking an assignment.
While developing this document, I ran this after each new segment below.

```{r check_for_missing_assignments }
check_for_missing_assignments <- hgcA_bin_metabolism_assignment %>%
  filter(bin_id %in% (hgcA_bin_data_final %>%
           filter(is.na(meta_code)) %>%
           select(bin_id) %>%
           unlist(use.names = FALSE)))
```

### BLI_hgcA_HMS_0011

All of the bins in BLI_hgcA_HMS_0011 have *nqr* and MHCs, but not clear what the TEAPs.
Each bin in the HMS has a hit to the pcc_porin HMM, so let's check the neighborhood:
- BLI21_assembly103_000000002038_7: Not near an MHC. Near a HemX protein (whatever that is) and a cytochrome c biogenesis protein.
- The other are the same.
All bins also have nrfAH, but no molybdopterin oxidoreductases.
Let's call this "RESP", given that the genomes likely have some capacity for respiration.

```{r meta_code_for_HMS_0011 }
hgcA_bin_data_final$meta_code[hgcA_bin_data_final$bin_id %in% (hgcA_bin_metabolism_assignment %>%
                                                                 filter(hms_id == "BLI_hgcA_HMS_0011") %>%
                                                                 select(bin_id) %>%
                                                                 unlist(use.names = FALSE))] <- "RESP"
```   

### BLI_hgcA_HMS_0006

The HMS BLI_hgcA_HMS_0006 is also missing an assignment.
All the bins have complete Rnf sequences and *nrfAH*.
Several have pcc_porin hits, which don't have annotations in KOFAMscan, and are close to HemX protein, glutamyl-tRNA reductase, and Fe-coproporphyrin III synthase.
Not sure if that means anything.
They all have at least 10 MHCs and *cyd*.
No complex I (neither *nuo* nor *nqr*) though and no MoOR hits.
Let's look more closely at the MHCs, see if there are hints in the KOFAMscan results.

```{r kofam_entries_for_MHCs_in_BLI_hgcA_HMS_0006 }
MHC_genes_in_HMS_0006 <- kofamscan_data %>%
  filter(gene_id %in% (MHC_raw_data %>%
                         filter(bin_id %in% (hgcA_bin_data %>%
                                               filter(hms_id == "BLI_hgcA_HMS_0006") %>%
                                               select(bin_id) %>%
                                               unlist(use.names = FALSE))) %>%
                         select(geneID) %>%
                         unlist(use.names = FALSE)))
```

The only ones that are present in the KOFAMscan data are the nitrite reductases.
Let's poke around, see what else is here.

```{r kofam_entries_for_BLI_hgcA_HMS_0006 }
kofam_genes_in_HMS_0006 <- kofamscan_data %>%
  filter(bin_id %in% (hgcA_bin_data %>%
                        filter(hms_id == "BLI_hgcA_HMS_0006") %>%
                        select(bin_id) %>%
                        unlist(use.names = FALSE)))
```

Most of these bins have NADH-quinone oxidoreductase subunit E and F.
These are directly adjacent to a NADP-reducing hydrogenase subunit HndD [EC:1.12.1.3], which is an FeFe hydrogenase.
Not particularly enlightening.

```{r check_BLI21_coassembly_anvio_bin_0181, include=FALSE}
test <- kofamscan_data %>%
  filter(bin_id %in% "BLI21_coassembly_anvio_bin_0181")
```

I doubt this is a respiratory organism. I think the nitrite reductases and the cytochrome bd oxidase are for oxidative stress protection.
The cytochrome bd oxidase does not pump protons, simply transfers charge, and has a very high O~2~ affinity, which makes it an efficient detoxifier.
Maybe that's what the remaining MHCs are for as well.
I'm going to call this one a "FERM", to distinguish from the "OB_FERM", but will aggregate their abundance for the paper figures.

```{r meta_code_for_HMS_0006 }
hgcA_bin_data_final$meta_code[hgcA_bin_data_final$bin_id %in% (hgcA_bin_metabolism_assignment %>%
                                                                 filter(hms_id == "BLI_hgcA_HMS_0006") %>%
                                                                 select(bin_id) %>%
                                                                 unlist(use.names = FALSE))] <- "FERM"
```   

### BLI_hgcA_HMS_0002

The bins in HMS BLI_hgcA_HMS_0002 also don't have an assignment.
This is clearly "RESP", as the bins have MHCs (including some with >= 10 heme-binding sites), *nuo*, MoORs, and cytochrome c and bd oxidases.
It's not clear what they respire, but for our purposes we'll stick with "RESP".

```{r meta_code_for_HMS_0002 }
hgcA_bin_data_final$meta_code[hgcA_bin_data_final$bin_id %in% (hgcA_bin_metabolism_assignment %>%
                                                                 filter(hms_id == "BLI_hgcA_HMS_0002") %>%
                                                                 select(bin_id) %>%
                                                                 unlist(use.names = FALSE))] <- "RESP"
```

### BLI_hgcA_HMS_0004 and BLI_hgcA_HMS_0016

Both of these HMS have one bin in them that was assigned a RESP based on the presence of a *narG* gene.
However, a quick glance at the remaining bins shows that they have *nuo*, MHCs, MoORs, oxidases, etc.
Naming them all as "RESP".

```{r meta_code_for_HMS_0004_and_0016 }
hgcA_bin_data_final$meta_code[hgcA_bin_data_final$bin_id %in% (hgcA_bin_metabolism_assignment %>%
                                                                 filter(hms_id %in% c("BLI_hgcA_HMS_0004", "BLI_hgcA_HMS_0016")) %>%
                                                                 select(bin_id) %>%
                                                                 unlist(use.names = FALSE))] <- "RESP"
```

### BLI_hgcA_HMS_0018

Four bins in the Kiritimatiellae class.

```{r kofam_entries_for_BLI_hgcA_HMS_0018 }
kofam_genes_in_HMS_0018 <- kofamscan_data %>%
  filter(bin_id %in% (hgcA_bin_data %>%
                        filter(hms_id == "BLI_hgcA_HMS_0018") %>%
                        select(bin_id) %>%
                        unlist(use.names = FALSE)))
```


All of these have *nuo* and the *rnf* chain.
They also all have *nrfAH*, *cydAB*, and a bunch of MHCs.
Definitely a respiratory group:

```{r meta_code_for_HMS_0018 }
hgcA_bin_data_final$meta_code[hgcA_bin_data_final$bin_id %in% (hgcA_bin_metabolism_assignment %>%
                                                                 filter(hms_id %in% c("BLI_hgcA_HMS_0018")) %>%
                                                                 select(bin_id) %>%
                                                                 unlist(use.names = FALSE))] <- "RESP"
```


### BLI_hgcA_HMS_0022

Next we have BLI_hgcA_HMS_0022, which in phylum Actinobacteriota, class Coriobacteriia, genus PHET01.
Two bins, both incomplete (BLI21_coassembly_anvio_bin_0181 -> 53%, 1.4Gbp long; BLI21_coassembly_dasTool_bin_0253 -> 62%, 1.8 Gbp long).
Both have a lot of MHCs, including 10 heme-binding sites.
BLI21_coassembly_dasTool_bin_0253 has *nuo*, but BLI21_coassembly_anvio_bin_0181 does not.
Could be that it didn't assemble in, given the incompleteness.
One MoOR.

```{r kofam_entries_for_BLI_hgcA_HMS_0022 }
kofam_genes_in_HMS_0022 <- kofamscan_data %>%
  filter(bin_id %in% (hgcA_bin_data %>%
                        filter(hms_id == "BLI_hgcA_HMS_0022") %>%
                        select(bin_id) %>%
                        unlist(use.names = FALSE)))
```

Let's see if the MHCs correspond to any KOFAM entry.

```{r kofam_entries_for_MHCs_in_BLI_hgcA_HMS_0022 }
MHC_genes_in_HMS_0022 <- kofamscan_data %>%
  filter(gene_id %in% (MHC_raw_data %>%
                         filter(bin_id %in% (hgcA_bin_data %>%
                                               filter(hms_id == "BLI_hgcA_HMS_0022") %>%
                                               select(bin_id) %>%
                                               unlist(use.names = FALSE))) %>%
                         select(geneID) %>%
                         unlist(use.names = FALSE)))
```

Wow, no KOFAMscan results for any of the MHCs in these bins.
Still, I feel pretty comfortable assigning this a RESP.

```{r meta_code_for_HMS_0022 }
hgcA_bin_data_final$meta_code[hgcA_bin_data_final$bin_id %in% (hgcA_bin_metabolism_assignment %>%
                                                                 filter(hms_id %in% c("BLI_hgcA_HMS_0022")) %>%
                                                                 select(bin_id) %>%
                                                                 unlist(use.names = FALSE))] <- "RESP"
```


### BLI_hgcA_bin_0001  and BLI_hgcA_bin_0004

These two bins are in the UBA1067 family of Kiritimatiellae.
The HMS BLI_hgcA_bin_0001 consists only of bin BLI20_coassembly_dasTool_bin_0110.
The HMS BLI_hgcA_bin_0004 consists only of bin BLI21_coassembly_anvio_bin_0184.
Just like BLI_hgcA_HMS_0006, they have complete Rnf sequences and *nrfAH*.
One of them (BLI_hgcA_bin_0001) has a *cyd* and two pcc_porin hits (no KOFAMscan annotations).
BLI20_coassembly_dasTool_bin_0110 has 9 MHCs and BLI21_coassembly_anvio_bin_0184 has 6 (all have less than 10 heme-binding sites).
No complex I (neither *nuo* nor *nqr*) though and no MoOR hits.
Let's look more closely at the MHCs, see if there are hints in the KOFAMscan results.

```{r kofam_entries_for_MHCs_in_BLI_hgcA_HMS_0001_0004 }
MHC_genes_in_HMSs_0001_0004 <- kofamscan_data %>%
  filter(gene_id %in% (MHC_raw_data %>%
                         filter(bin_id %in% (hgcA_bin_data %>%
                                               filter(hms_id %in% c("BLI_hgcA_bin_0001", "BLI_hgcA_bin_0004")) %>%
                                               select(bin_id) %>%
                                               unlist(use.names = FALSE))) %>%
                         select(geneID) %>%
                         unlist(use.names = FALSE)))
```

The only ones that are present in the KOFAMscan data are the nitrite reductases.
Let's poke around, see what else is here.

```{r kofam_entries_for_BLI_hgcA_HMS_0001_004 }
kofam_genes_in_HMS_0001_0004 <- kofamscan_data %>%
  filter(bin_id %in% (hgcA_bin_data %>%
                        filter(hms_id %in% c("BLI_hgcA_bin_0001", "BLI_hgcA_bin_0004")) %>%
                        select(bin_id) %>%
                        unlist(use.names = FALSE)))
```

I doubt these are respiratory organisms, comparable to how we described BLI_hgcA_bin_0006.

```{r meta_code_for_HMS_0001_004 }
hgcA_bin_data_final$meta_code[hgcA_bin_data_final$bin_id %in% (hgcA_bin_metabolism_assignment %>%
                                                                 filter(hms_id %in% c("BLI_hgcA_bin_0001", "BLI_hgcA_bin_0004")) %>%
                                                                 select(bin_id) %>%
                                                                 unlist(use.names = FALSE))] <- "FERM"
```   

ASSIGNMENTS COMPLETE.

```{r check_for_missing_assignments_last }
check_for_missing_assignments_last <- hgcA_bin_metabolism_assignment %>%
  filter(bin_id %in% (hgcA_bin_data_final %>%
           filter(is.na(meta_code)) %>%
           select(bin_id) %>%
           unlist(use.names = FALSE)))
```


# Save out assignment data

```{r save_out_data }
saveRDS(hgcA_bin_data_final,file = 'dataEdited/bin_based_analyses/bin_data_aggregate_2_metabolism.rds')
```
